### 2022-02-27 chapters/layouts.tex

\chapter{Layouts}

\begin{figure}
\listdiagram
\caption{The \LaTeX{} parameters which define typesetting and layout of lists.} 
\label{fig:layout-list}
\end{figure}

% \begin{figure}
% \oddpagelayoutfalse
% \pagediagram
% \caption{Left-hand page layout parameters}
% \label{fig:layout-left-page}
% \end{figure}

% \begin{figure}
% \oddpagelayouttrue
% \pagediagram
% \caption{Right-hand page layout parameters}
% \label{fig:layout-right-page}
% \end{figure}

\begin{figure}
\oddpagelayoutfalse
\twocolumnlayoutfalse
\stockdiagram
\caption{Left-hand page major layout parameters for \file{memoir} class}
\label{fig:layout-left-memoir}
\end{figure}

\begin{figure}
\oddpagelayouttrue
\twocolumnlayoutfalse
\stockdiagram
\caption{Right-hand page major layout parameters for \file{memoir} class}
\label{fig:layout-right-memoir}
\end{figure}