### 2022-02-27 preamble/static.tex

%!TEX root = ../Thesis.tex

% This information is used in titlepage, colophon, preface and hyperref setup (pdf metainfo), and other options.

%\def\thesistypeabbr{B.Eng.}
%\def\thesistype    {Bachelor of Engineering}
%\def\thesistypeabbr{B.Sc.Eng.}
%\def\thesistype    {Bachelor of Science in Engineering}
%\def\thesistypeabbr{M.Sc.}
%\def\thesistype    {Master of Science in Engineering}
\def\thesistypeabbr{Ph.D.}
\def\thesistype    {Doctor of Philosophy}

\def\thesisdepshort{DTU Compute}
\def\thesisdep     {Department of Applied Mathematics and Computer Science}

\def\thesisauthor  {Jakob Drachmann Havtorn}
\def\thesistitle   {Uncertainty Quantification in Machine Learning}
\def\thesissubtitle{}
\def\thesislocation{Copenhagen}
\def\thesisyear    {2023}
\def\thesismonth   {August}
\def\thesisday     {31}

\def\papersize     {b5paper} % Final papersize (b5paper/a4paper), recommended papersize for DTU Compute is b5paper
\def\showtrims     {false}   % Print on larger paper than \papersize and show trim marks (true/false)?
\def\showframe     {false}   % Show frame of writeable area, marginparsep and marginpar (true/false)?

\def\showtodos     {true}    % Show todos (true/false)?
\def\confidential  {false}   % Confidential thesis (true/false)?