### 2022-02-27 prefrontmatter/colophon.tex

%!TEX root = ../Thesis.tex

\thispagestyle{empty} % No page numbers
\frieze{0}{0.5\paperheight}

\vspace*{\fill}
% \vspace*{0.15\textheight}
\small\sffamily
\noindent
\textsc{\textbf{DTU Compute}}\\
\textsc{\textbf{Department of Applied Mathematics and Computer Science}}\\
\textsc{\textbf{Technical University of Denmark}}
\smallskip\\
Matematiktorvet, Building 303B\\
2800 Kongens Lyngby, Denmark\\
% Phone +45 4525 3031\\
compute@compute.dtu.dk\\
www.compute.dtu.dk
\bigskip\\
\noindent
\textsc{\textbf{Corti}}\\
\textsc{\textbf{Research Department}}
\smallskip\\
Store Strandstræde 21. 4.tv\\
1255 København K, Denmark\\
info@corti.ai\\
www.corti.ai\\

\normalsize
\normalfont