### 2023-01-06 frontmatter/abstract-da.tex
Fix name highlighting and update font

%!TEX root = ../thesis.tex
\chapter[resumé (danish)]{Resumé (Danish)}
\blindtext[5]