### 2023-01-06 frontmatter/acknowledgements.tex

%!TEX root = ../thesis.tex
\chapter[acknowledgements]{Acknowledgements}
First of all, I want to thank all my supervisors...

\vspace*{\fill}
\noindent \textit{The project was funded by Corti and Innovation Fund Denmark through industrial PhD grant no.\@ 0153-00167B).}