### 2023-01-06 prefrontmatter/colophon.tex

%!TEX root = ../thesis.tex

\thispagestyle{empty} % No page numbers
\frieze{0}{0.5\paperheight}

\vspace*{\fill}
% \vspace*{0.15\textheight}
\small\sffamily
\noindent
\textsc{\textbf{Technical University of Denmark}}\\
\textsc{\textbf{Department of Applied Mathematics and Computer Science}}
% \textsc{\textbf{DTU Compute}}
\smallskip\\
\href{https://goo.gl/maps/Dx6rNjGT9Ad5xKgZ6}{%
Richard Petersens Plads, Building 324\\
2800 Kongens Lyngby, Denmark}\\
% Phone +45 4525 3031\\
\href{mailto:compute@compute.dtu.dk}{compute@compute.dtu.dk}\\
\href{www.compute.dtu.dk}{www.compute.dtu.dk}
\bigskip\\
\noindent
\textsc{\textbf{Corti}}\\
\textsc{\textbf{Research Department}}
\smallskip\\
\href{https://goo.gl/maps/X7en4p8vaHkfVuyJ7}{%
Store Strandstræde 21. 4.tv\\
1255 København K, Denmark}\\
\href{mailto:info@corti.ai}{info@corti.ai}\\
\href{www.corti.ai}{www.corti.ai}\\

\normalsize
\normalfont