### 2023-03-04 chapters/paper_review/02_three_waves.tex

%!TEX root = ../thesis.tex

\section{Historical Context of Representation Learning}

\label{sec:thirdwave}

% {\color{blue} Reviewer: KL, KK, Daniel }\\
In this section we present the historical background of the current surge in
self-supervised representation learning methods in the context of two previous
waves of research work in the 1990s and 2000s. The discussed approaches go
beyond speech to describe the overall landscape of machine learning development during the
past few decades.

%%KK: if you adopt the paragraph below I suggest renaming this as "Features based on generative models" or something
\subsection{Clustering and mixture models}
% {\color{blue} Tara}\\

%\kl{to people who are familiar with modern rep learning work, this first wave may not seem like rep learning at all.  readers might also wonder why we are talking about what look like standard supervised ASR in the 3rd paragraph below.  one option is to describe it is as unsupervised learning of latent variables (Gaussian components, HMM states) within a supervised model. (something like that -- I haven't thought about the wording carefully)}
%KK: suggesting the paragraph below for inclusion here
%The initial wave of speech representation {\em learning} focused on extracting speech features from generative models. The main objective was
%to render  the knowledgpe learned by generative models accessible to
%discriminative downstream classifiers, or to map variable-length sequences into  fixed-length representations.
%The generative model, typically a global Gaussian mixture model (GMM) or a Hidden Markov Model (HMM) was trained by
%by maximizing the likelihood of the data given the
%model, which could be done in either an unsupervised or a supervised way. Feature vectors were then derived from the parameters of the trained model. In
%the case of {\em Fisher vectors}, the features were the normalized gradients of the log-likelihood
%with respect to the model parameters (mixture weights, means, variances) of the Gaussian mixtures.
%An extension of this approach (likelihood ratio score space) used the derivative of the log likelihood ratio of	two models,
%e.g., a background model and a foreground model.
%Examples of their use in speech processing include speech recognition \cite{smith01,venkata03} and speaker recognition \cite{wan03}.
%Subsequent techniques in speaker and language verification \cite{dehak11a,dehak11b} similarly extracted parameters (concatenated means) from trained background GMMs as representations that were then %combined with low-rank projections of speaker/session or language specific vectors.

Initial research in learning latent speech and audio representations involved
simple models in which the training data likelihood was optimized directly
or via the expectation--maximization (EM) algorithm.

Early work used simple clustering methods. For example, in work such
as \cite{Rabiner1979,Wilpon1985}, word patterns were clustered
semi-automatically using techniques such as k-means, after which isolated words
were recognized by finding the training cluster closest to 
  the test data.   % AMH: check

Through time, modeling techniques improved such that subword units were
represented by Gaussian mixture models (GMMs) \cite{Gauvain1994}, which facilitated
the modeling of more variability in the input data. GMMs were first built for
context-independent phonemes; state-clustering 
algorithms~\cite{Young1994} then resulted in GMMs for context-dependent phonemes. 
Each latent component of these mixture models acted as a template of a
prototypical speech frame, 
% leading to their deficiency to deal with 
  making it difficult to handle             % AMH: check
large volumes of data with diverse characteristics. 
Furthermore, dynamical models like hidden Markov models (HMMs)~\cite{Bahl1986}
allowed for the processing of continuous speech rather than just isolated word
recognition. These generative GMM and HMM models were trained by maximizing the
likelihood of data given the model, which could be accomplished in either an
unsupervised or a supervised manner.

Another line of research focused on extracting speech features from generative
models. The main objective here was to render the knowledge learned by generative
models accessible to discriminative downstream classifiers, or to map
variable-length sequences to fixed-length representations. Feature vectors
were derived from the parameters of trained GMM models. In the case of {\em
Fisher vectors}, the features were the normalized gradients of the
log-likelihood with respect to the model parameters (mixture weights, means, and
variances) of the Gaussian mixtures. An extension of this approach (likelihood
ratio score space) used the derivative of the log-likelihood ratio of two
models, e.g., a background model and a foreground model. Examples of their use
in speech processing include speech recognition~\cite{smith01,venkata03} and
speaker recognition~\cite{wan03}. Subsequent techniques in speaker and language
verification~\cite{dehak11a,dehak11b} similarly extracted parameters
(concatenated means) from trained background GMMs as representations that were
then combined with low-rank projections of speaker/session- or language-specific
vectors.

\subsection{Stacked neural models}
\label{subsec:stack}

%{\color{blue} Tara}\\
More recently, representation learning has seen a shift of focus towards neural models,
which, compared to GMMs and HMMs, offer distributed representations with more
capacity to model diverse input signals into efficient latent binary codes. 
Examples of early techniques include restricted Boltzmann machines
(RBM)~\cite{hinton_2006}, denoising autoencoders~\cite{DAE}, noise contrastive
estimation (NCE)~\cite{gutmann2012noise}, sparse coding~\cite{Olshausen1996,
sparse_lee, sivaram2010sparse}, and energy-based methods~\cite{Ranzato2007}.
Many of these techniques have also been applied to CV and NLP problems, which
provided inspiration for their application to speech.
%Examples of techniques include Restricted Boltzmann Machine (RBM) \cite{hinton2012practical,Srivastava2013}, Neural Auto Encoders \cite{Lecun2012}, Noise Contrastive Estimation (NCE) \cite{chen2015recurrent},  Sparse Coding \cite{sivaram2010sparse} and Energy-based method \cite{Ranzato2007}. Note that many of these techniques were also tried for CV and NLP problems, which provided inspiration to try them for ASR as well.

Higher-capacity neural models were achieved by stacking several neural network
layers to build progressively higher-level concept representations.
However, these deeper networks also increased the training complexities. For
example, approximate training methods such as contrastive divergence
\cite{hinton_cd_2002} were a practical technique to 
% make       RBM training more efficient. 
  streamline RBM training.                   % AMH: check
Furthermore, deep networks had non-convex objective functions, which
often resulted in long training times compared to GMMs, which are trained
using full batches instead of mini-batch learning.

% http://proceedings.mlr.press/v2/ranzato07a/ranzato07a.pdf, 
%https://auai.org/uai2013/prints/papers/166.pdf, % https://www.cs.cmu.edu/~rsalakhu/papers/neco_DBM.pdf
% problems with this trend: approximate gradient, failure modes, long training time, no clear wins in downstream applications.

\subsection{Learning through pretext task optimization}

% {\color{blue} Hung-yi + Abdo}\\

%Note on 12/09: We will mainly describe the speech part but link to CV and NLP. 
%discuss instance classification, contrastive losses,\\

%Currently, the network learns a function that maps input to desired representations by solving a \textit{pretext} task.
%After pre-training, learned representation models could be applied to downstream tasks through feature-based representation extraction or fine-tuning as part of the downstream model.  
%By optimizing carefully designed pretext tasks, neural models can capture latent representations which generalize well across a wide range of downstream applications, usually under few-shot or even zero-shot conditions. 
A more recent trend is learning networks that map the input to desired
representations by solving a \textit{pretext} task. Such studies have several
characteristics:
(1)~All layers are trained end-to-end to optimize a single pretext task instead
of relying on layer-wise pre-training
(2)~Past stacked networks typically had only a few layers, but 
very deep networks with more than ten layers are now common.
(3)~It is common to evaluate a representation model on a wide range of tasks.
For example, in NLP, a representation model is usually assessed on GLUE,
which comprises nine tasks~\cite{GLUE}, whereas in speech, a representation model can be
evaluated on SUPERB, which comprises ten tasks~\cite{yang21c_interspeech}, 
as described in detail in \cref{sec:benchmark}.

The cornerstone of this third wave is the design of a pretext task, which
allows the model to efficiently leverage knowledge from unlabeled data.
The pretext task should be challenging enough for the model to learn high-level
abstract representations and 
% not be too amiable to exploit low-level shortcuts.
  not be so easy as to encourage the exploitation of low-level shortcuts.  % AMH: check
Early breakthroughs included end-to-end learning of deep neural architectures
via pretext tasks for restoring the true color of black-and-white
images~\cite{colorizing}, joint learning of latent representations and their
cluster assignments~\cite{deepcluster}, and the prediction of the relative positions of
image patches~\cite{context_pred}. Other popular approaches include variational
autoencoders (VAEs)~\cite{vae, rezende2014stochastic}. While typical autoencoders learn data
representations using unsupervised objectives by reconstructing the input
after passing it through an information bottleneck, VAEs estimate a neural model of a probability density function (pdf) that approximates the unknown “true” distribution of the observed data, for which we only have access to independently identically distributed (iid) samples. It is also important to mention dynamical VAEs \cite{Girin2021}, which is an extension of VAE for sequential data such as speech.

In the SSL context, a pretext task related to autoencoding is to generate
an object from its partial information. Such tasks are widely used in NLP, for
example, using 
the previous tokens 
in a sentence to predict the next token such as in
ELMo~\cite{peters2018deep}, the GPT
series~\cite{alex2019GPT2}, and %brown2020gpt3 alex2018GPT,
Megatron~\cite{shoeybi2020megatronlm}, or predicting the masked tokens in a
sentence such as with the bidirectional encoder representations from Transformers
(BERT) series~\cite{jacob2019BERT,Liu2019RoBERTa}. %,albert
%ELMo~\cite{peters2018deep}, GPT series~\cite{alex2018GPT,alex2019GPT2,brown2020gpt3}, Megatron~\cite{shoeybi2020megatronlm} are gigantic LM which predict the next token. 
%BERT~\cite{jacob2019BERT}, RoBERTa~\cite{Liu2019RoBERTa}, ALBERT~\cite{albert} masked the input token, and the models are learned to predict the masked tokens. 
Another common pretext task in the third wave is contrastive
learning~\cite{oord2018representation}, in which a model learns to identify a
target instance from a set of negative samples. 
This approach has become especially popular in the
CV context~\cite{pmlr-v119-chen20j,he2020momentum,chen2020improved,SwAV}.
In this survey, we will mainly focus on techniques for pretext task
optimization for speech processing, and discuss these techniques in detail
in \cref{sec:approach}.

\subsection{Other related work}
A closely related area of research that is not covered in this review is
semi-supervised pre-training methods such as pseudo-labeling (that is,
self-training). Pseudo-labeling (PL) relies on a supervised teacher model to
label a large volume of speech-only data, which is then used to augment the
initial labeled data to train a student model~\cite{Kemp1999, csl01_limsi,
ma_bbn_06, hari_1mhour}. PL has been successful and widely adopted in the
speech community since the 1990s. Other proposed variations of PL include
augmenting speech-only data with noise to improve robustness, iterating
over the PL process to improve teacher labeling quality, and training student
models with more parameters than their original teachers to capture the
complexities in vastly larger speech-only data~\cite{park2020improved,
xu2020iterative, xiao_scaling_2021}. 
Both SSL and PL leverage unlabeled speech-only data.
One distinguishing factor in PL is the utilization of supervised data for a
specific task during model pre-training, which limits the model's focus to
a single (or at best a few) downstream tasks. 
SSL, in turn, is an attempt to learn task-agnostic representations to benefit
a wide range of tasks.

Transfer learning (TL) is another closely related area of research for
pre-training speech models. TL transfers knowledge captured by models
trained on one task to different but related tasks~\cite{caruana1997multitask}. 
The past few decades have seen active research on TL and its
extension to multitask learning for more general representations. 
Multilingual and cross-lingual supervised models have proven superior in
low-resource speech recognition tasks~\cite{Cui2015MultilingualRF}.
SSL can be regarded as a type of TL because knowledge learned from pre-training
is used for different downstream tasks.
This survey paper focuses on SSL, and not all TL technologies for speech. 
One survey indeed addresses TL for speech processing~\cite{survey_speech_TL}
but does not include current SSL technologies for speech.

%Then contrastive loss is applied~\cite{oord2018representation}, in which the model learns to classify the positive instances from a set of unrelated negative samples. SimCLR~\cite{pmlr-v119-chen20j}, MoCo~\cite{he2020momentum,chen2020improved} and SwAV~\cite{SwAV} are representative ones. 
%These approaches learn representations from visual inputs by maximizing agreement between differently augmented views of the same sample while minimizing the agreement between different samples. 
%Then there is a series of work trying to remove the requirement of negative examples. 
%BYOL~\cite{BYOL} trains a student network to predict the representations of the teacher network.  The parameter of the teacher network is the exponential moving average of the previous student network. In such a scenario, negative examples are not required. Then SimSiam~\cite{Chen_2021_CVPR} improved BYOL by removing the moving average and found that good representation can still be learned in this way.
%In Barlow twins~\cite{BarlowTwins}, the cross-correlation matrix between the representations of two identical encoders fed with distorted versions of a sample is measured. The encoder learns to make the cross-correlation matrix as close to the identity matrix as possible.
%VICReg~\cite{bardes2021vicreg}

%The 3rd wave begins from NLP and CV communities. %Is it true?
%Below the progress of self-supervised learning in CV  and NLP are briefly summarized. %(However, we have to consider how to connect this part with other parts.)

%In CV, some earlier approaches make some pretexts to train instance classifiers, and the last hidden layers of the classifiers are used as the representations in the downstream tasks. Below are some representative approaches:
%\begin{itemize}
%\item Position~\cite{Doersch_2015_ICCV}: predicting the relative position between two random patches from one image. A model needs to understand the spatial context of objects to tell the relative position between parts.
%\item Jigsaw~\cite{noroozi2016unsupervised}: The model is trained to place shuffled patches back to the original locations.
%\item Rotation~\cite{gidaris2018unsupervised}: The model is trained to predict which rotation has been applied.
%\end{itemize}

%How to deal with negative examples is critical in this series. SimCLR~\cite{pmlr-v119-chen20j} uses a large batch size to obtain a lot of negative examples, while MoCo uses a momentum encoder to make the negative samples’ representation consistent.

%In ELECTRA~\cite{clark2020electra}, given a sentence with some tokens replaced, the ELECTRA model predicts whether a token is generated or not. 
%XLNet~\cite{XLNet}
%MASS~\cite{MASS}, BART~\cite{lewis-etal-2020-bart}, and T5~\cite{T5} are seq2seq models. The input is disturbed by different approaches, and the seq2seq model learns to reconstruct the disturbed input. %(There are speech T5, should we mention this work?)
%The contrastive approaches have been applied at the sentence level in NLP~\cite{gao-etal-2021-simcse,kim-etal-2021-self,yan-etal-2021-consert,giorgi-etal-2021-declutr}.

%Turing NLG only has webpage: https://www.microsoft.com/en-us/research/blog/turing-nlg-a-17-billion-parameter-language-model-by-microsoft/
%I don't know how to cite it.

%ref: http://speech.ee.ntu.edu.tw/~tlkagk/courses/DLHLP20/BERT%20train%20(v8).pdf