### 2023-03-04 chapters/paper_review/model.tex

%!TEX root = ../thesis.tex

\begin{table*}[h!]
\caption{
The models are sorted according to publication date of the first version on arXiv which might differ from the year indicated by the citation. 
\textbc{{reg}} means that the model uses a regression loss, 
\textbc{{clf}} that it uses a classification loss,
\textbc{{con}} that it uses a contrastive loss, 
\textbc{{msk}} that it uses masking to define the target region, 
\textbc{{ltg}} that it uses a learned target, 
\textbc{{qtz}} that it uses a quantized target, and
\textbc{{emb}} that it embeds a variable length segment.
%Lee: I think contrastive always have learned target. Am I correct???
}
\label{tab:model-taxonomy}
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{ l l | c c c | c c c | c } 
\toprule
\multicolumn{2}{c}{} & 
\multicolumn{3}{c}{\small\textsc{Loss}} & 
\multicolumn{3}{c}{\small\textsc{Target}} &
 \\
\textbc{{model}} &
\textbc{{pub. date}} &
\textbc{{reg}} & 
\textbc{{clf}} & 
\textbc{{con}} &
\textbc{{msk}} &
\textbc{{ltg}} &  
\textbc{{qtz}} & 
\textbc{{emb}} \\
\midrule
%                                                         REG      CLF      CON      MSK      LTG      QTZ      EMB

Audio Word2vec \cite{Chung2016AudioWord2Vec}      & 2016 Mar. & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \cmark \\

VQ-VAE \cite{vqvae} & 2017 Nov. & \cmark & \xmark & \xmark & \xmark & \xmark & \cmark & \xmark \\ %2 Nov 2017

Speech2Vec \cite{chung2018speech2vec}     & 2018 Mar. & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \cmark \\
Unspeech \cite{milde2018unspeech}         & 2018 Apr. & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark & \cmark \\

CPC \cite{oord2018representation}         & 2018 Jul. & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark \\

PASE \cite{pascual2019learning}                   & 2019 Apr. & \cmark & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark \\ % 6/4

APC \cite{chung2019unsupervised}          & 2019 Oct. & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\ % 5/4

wav2vec \cite{schneider2019wav2vec}       & 2019 Apr. & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark \\ % 11/4

vq-wav2vec \cite{Baevski2020vq-wav2vec} & 2019 Oct. & \xmark & \xmark & \cmark & \xmark & \cmark & \cmark & \xmark \\ %12 Oct 2019 %Lee: vq-wav2vec is CON or CLF??? %Lee: Discrete BERT use masking, but vq-wav2vec does not It is CON, not CLF.

Discrete BERT \cite{Baevski2020vq-wav2vec,baevski2019effectiveness} & 2019 Oct. & \xmark & \cmark & \cmark & \cmark & \cmark & \cmark & \xmark \\ %12 Oct 2019 %10 Nov 2019 %In the paper of vq-wav2vec, it also includes the approach of Discrete BERT. I know this is tricky ....

Mockingjay  \cite{liu2020mockingjay}      & 2019 Oct. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark \\ % 25/10-19

DeCoAR \cite{ling2020deep}                & 2019 Dec. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark \\ % 3/12-19

PASE+ \cite{ravanelli2020multi}                & 2020 Jan. & \cmark & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark \\ % 25/1 % I think contrastive always have learned target.

Bidir CPC \cite{kawakami2020learning} & 2020 Jan. & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark \\% 29/1

Modified CPC \cite{riviere2020unsupervised} & 2020 Feb. & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark \\ % 7/2

VQ-APC \cite{chung20e_interspeech}  & 2020 May  & \cmark & \xmark & \xmark & \xmark & \xmark & \cmark & \xmark \\ %17/05

AALBERT \cite{chi2020audio}  & 2019 May & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark \\ %18 May 2020

speech-XLNet \cite{song20d_interspeech} & 2019 Oct. & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\%23 Oct 2019 %Lee: It is speech version of XLNet. I am not sure we should consider this approach has mask or not.

wav2vec 2.0 \cite{baevski2020wav2vec}     & 2020 Jun. & \xmark & \xmark & \cmark & \cmark & \cmark & \cmark & \xmark \\ % 20/6
%Lee: In my point of view, wav2vec 2.0 has learned target. The target is from the CNN encoder.

MPE \cite{liu2020masked} & 2020 May & \cmark & \xmark & \xmark & \cmark & \cmark & \xmark & \xmark \\ % 25 May 2020 %Lee: I think it has learned target. But please check.

TERA \cite{liu2021tera}                   & 2020 Jul. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark \\ % 12/7

Speech SimCLR \cite{SpeechSimCLR} &  2020 Oct. & \cmark & \xmark & \cmark & \cmark & \cmark & \xmark & \xmark \\ %Lee: It is TERA+ SIMCLR. I considered it has learned target because contrastive always have learned target.
%27 Oct 2020

NPC \cite{liu21l_interspeech}                     & 2020 Nov. & \cmark & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark \\ % 1/11

DeCoAR 2.0 \cite{ling2020decoar}          & 2020 Dec. & \cmark & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark \\ % 11/12 %Lee:  DeCoAR 2.0 has VQ layer although it is not used as learning target.

wav2vec-C \cite{sadhu21_interspeech} & 2021 Mar. & \cmark & \xmark & \cmark & \cmark & \cmark & \cmark & \xmark \\ %9 Mar 2021 %Lee:  consistency loss is reconstruction

HuBERT \cite{hsu2021hubert}               & 2021 Jun. & \xmark & \cmark & \xmark & \cmark & \cmark & \cmark & \xmark \\ % 14/6

W2v-bert \cite{w2vbert} & 2021 Aug & \xmark & \cmark & \cmark & \cmark & \cmark & \cmark & \xmark \\
%7 Aug 2021

pMPC \cite{yue2021pMPC}  & 2021 Aug. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark \\ %30 August – 3 September, 2021 (I do not find arXiv version.)

WavLM \cite{chen2021wavlm} & 2021 Oct. & \xmark & \cmark & \xmark & \cmark & \cmark & \cmark & \xmark \\ %26 Oct 2021

data2vec \cite{data2vec}     & 2022 Jan. & \cmark & \xmark & \xmark & \cmark & \cmark & \xmark & \xmark \\ % 20/1

BEST-RQ \cite{BEST-RQ} & 2022 Feb.  & \xmark & \cmark & \xmark & \cmark & \xmark & \cmark & \xmark \\ %Lee: no learned target. The target is random ...
% 3 Feb 2022

\bottomrule
\end{tabular}
\end{center}
\end{table*}

\begin{comment}
\begin{table*}[h!]
\caption{
The models are sorted according to the publication date of the first version on arXiv, which might differ from the year indicated by the citation. 
\textbc{{reg}} means that the model uses a regression loss, 
\textbc{{clf}} that it uses a classification loss,
\textbc{{con}} that it uses a contrastive loss, 
\textbc{{msk}} that it uses masking to define the target region, 
\textbc{{ltg}} that it uses a learned target, %lasseborgholt: We could consider using the inverse of this attribute. I.e., that the model uses the input as taget and belongs in the generative category.
\textbc{{qtz}} that it uses a quantized target, and
\textbc{{emb}} that it embeds a variable-length segment
}
\label{tab:model-taxonomy}
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{ l l | c c c | c c c | c } 
\toprule
\multicolumn{2}{c}{} & 
\multicolumn{3}{c}{\small\textsc{Loss}} & 
\multicolumn{3}{c}{\small\textsc{Target}} &
 \\
\textbc{{model}} &
\textbc{{pub. date}} &
\textbc{{reg}} & 
\textbc{{clf}} & 
\textbc{{con}} &
\textbc{{msk}} &
\textbc{{ltg}} &  
\textbc{{qtz}} & 
\textbc{{emb}} \\
\midrule
%                                                         REG      CLF      CON      MSK      LTG      QTZ      EMB  
Audio Word2vec \cite{Chung2016AudioWord2Vec}      & 2016 Mar. & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \cmark \\
Speech2Vec \cite{chung2018speech2vec}     & 2018 Mar. & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \cmark \\
% Unspeech \cite{milde2018unspeech}         & 2018 Apr. & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark & \cmark \\
CPC \cite{oord2018representation}         & 2018 Jul. & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark \\

PASE \cite{pascual2019learning}                   & 2019 Apr. & \cmark & \xmark & \cmark & \xmark & \xmark & \xmark & \xmark \\ % 6/4

APC \cite{chung2019unsupervised}          & 2019 Oct. & \cmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\ % 5/4
wav2vec \cite{schneider2019wav2vec}       & 2019 Apr. & \xmark & \xmark & \cmark & \xmark & \cmark & \xmark & \xmark \\ % 11/4
%VQ-APC \cite{chung2020vector}             & 2020 May  & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\ 
Mockingjay  \cite{liu2020mockingjay}      & 2019 Oct. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark \\ % 25/10-19
%DeCoAR \cite{ling2020deep}                & 2019 Dec. & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\ % 3/12-19
%PASE+ \cite{Ravanelli2020}                & 2020 Jan. & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\ % 25/1
wav2vec 2.0 \cite{baevski2020wav2vec}     & 2020 Jun. & \xmark & \xmark & \cmark & \cmark & \xmark & \cmark & \xmark \\ % 20/6
%TERA \cite{liu2021tera}                   & 2021 Jul. & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark \\ % 12/7

NPC \cite{liu21l_interspeech}             & 2020 Nov. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark \\ % 1/11

DeCoAR 2.0 \cite{ling2020decoar}          & 2020 Dec. & \cmark & \xmark & \xmark & \cmark & \xmark & \xmark & \xmark \\ % 11/12
HuBERT \cite{hsu2021hubert}               & 2021 Jun. & \xmark & \cmark & \xmark & \cmark & \cmark & \cmark & \xmark \\ % 14/6
data2vec \cite{data2vec}     & 2022 Jan. & \cmark & \xmark & \xmark & \cmark & \cmark & \xmark & \xmark \\ % 20/1
\bottomrule
\end{tabular}
\end{center}
\end{table*}

\end{comment}