### 2023-03-04 chapters/paper_review/pretext.tex

%!TEX root = ../thesis.tex

\begin{table*}[!htb]
    \centering
    \caption{
    A summary of the approaches in the three categories of
    self-supervised learning. 
    Column~(a) lists the names of the models and related references, 
    column~(b) defines the model input, 
    column~(c) defines any corruption of the input or hidden representation, and
    column~(d) defines the target of the pretext task; the pretext task itself
    is described by the overall model category and the main text.
    $X=\{x_1,x_2,...,x_T\}$ is the input sequence in which $x_t$ can be an
    acoustic feature vector (e.g., MFCC, filterbank, or spectrogram features)
    or a waveform sample. 
    $X_{[t_1:t_2]}$ represents $\{x_{t_1},x_{t_1+1},...,x_{t_2}\}$.
    $X_{-[t_1:t_2]}$ represents $X$ in which the segment
    $X_{[t_1:t_2]}=\{x_{t_1},x_{t_1+1},...,x_{t_2}\}$ is masked.
    $x_{t}^{i}$ represents the $i$-th dimension of $x_t$.
    If $x_t$ is a frame in a spectrogram, then the $i$-th dimension corresponds
    to a specific frequency bin.
    $X^{-[f,f+j]}$ refers to a spectrogram $X$ which is masked along the frequency
    axis from the $f$-th to $(f+j)$-th bin. 
    % $X_{-[t:t+k]}^{-[f,f+j]}$ refers to a spectrogram $X$ but masked along both time and frequency dimensions. 
    % $X^*$ is a temporally permuted version of $X$, that is, the $x_t$ are randomly shuffled to form $X^*$. 
    We indicate random temporal permutation of a sequence by indexing it with
    the set $\mathcal{P}_t\triangleq\textsc{permute}([0,t])$, where
    $\textsc{permute}(\cdot)$ returns a permutation of the given list. 
    We indicate data augmentation (e.g., reverberation) by the function
    $\textsc{augment}(\cdot)$. Subscripts indicate different augmentations. 
    % $X^\prime$ is an augmented version of $X$ (e.g., $X$ with reverberation), while $X^{\prime\prime}$ is $X$ adding distortion different from  $X^\prime$.
    $Z$ represents a localized latent representation sequence of $X$. %, and $Z^\prime$ and $Z^{\prime\prime}$ are the latent representation sequences of  $X^\prime$ and $X^{\prime\prime}$, respectively. 
    % $Z$ represents a localized latent representation sequence and $Z^\prime$ is the latent representation sequence of $X^\prime$. 
    $Z^{(l)}$ is $Z$ at the $l$-th layer of the model used to compute it.
    $\bar{H}$ is the contextualized sequence $H$ obtained from an exponential
    moving average (EMA) of the model undergoing training with no masking
    applied.
    $Q$ represents a sequence of quantized learned representations, and $C$ is
    a sequence of discrete cluster IDs.
    For contrastive models, we specify only positive targets.
    }
    % \begin{minipage}{\textwidth}  
\centering
\renewcommand*\arraystretch{1.2}{
\begin{tabular}{l|c|c|c}
    \toprule
    \textbf{Model} (a) & \textbf{Input} (b) & \textbf{Corruption} (c) & \textbf{Target} (d) \\
    \midrule
    \midrule
    \multicolumn{4}{c}{\textsc{Generative models}} \\
    \midrule
    \midrule
    Audio Word2vec~\cite{Chung2016AudioWord2Vec}, VQ-VAE~\cite{vqvae}     & $X$ &   \textsc{-}    &  $X$  \\ %It has VQ as extra constraint.
    \midrule  
    Speech2Vec~\cite{chung2018speech2vec}, Audio2Vec~\cite{tagliasacchi2020pre} - skip-gram    & $X_{[t_1,t_2]}$  &     \textsc{-}   &    $X_{[t_0,t_1]}$,$X_{[t_2,t_3]}$     \\
    \midrule 
    Speech2Vec~\cite{chung2018speech2vec}, Audio2Vec~\cite{tagliasacchi2020pre} - cbow    & $X_{[t_0,t_1]}$,$X_{[t_2,t_3]}$   &     \textsc{-}   &    $X_{[t_1,t_2]}$   \\
    \midrule 
    PASE~\cite{pascual2019learning}, PASE+~\cite{ravanelli2020multi}\footnote{PASE uses multiple pretext tasks, but the authors find that reconstruction is most important.}       & $X$ &  \textsc{-}   &  Different modalities of $X$  \\
    \midrule 
    APC~\cite{chung2019unsupervised,chung20e_interspeech}         & $X_{[1,t]}$   & \textsc{-}              & $x_{t+c},\, c\geq1$    \\ %VQ-APC is the same as APC
    \midrule
    Speech-XLNet \cite{song20d_interspeech}     & \multicolumn{2}{c|}{$X_{\mathcal{P}_{t}}$}   &     $x_{i\sim\mathcal{P}^c_{t}}$  \\ %Lee: Use * to represent the permutation, hope it is not strange.
    \midrule  
    DeCoAR~\cite{ling2020deep}     & $X_{[1,t-1]}, X_{[t+k+1,T]}$ & \textsc{-} & $X_{[t,t+k]}$   \\
    \midrule
    Mockingjay~\cite{liu2020mockingjay}, Audio ALBERT~\cite{chi2020audio}, DeCoAR 2.0~\cite{ling2020decoar}   & \multicolumn{2}{c|}{$X_{-[t,t+k]}$}   & $X_{[t,t+k]}$    \\
    \midrule 
    TERA~\cite{liu2021tera}, BMR~\cite{wang2020unsupervised}  & \multicolumn{2}{c|}{$X_{-[t,t+k]}^{-[f,f+j]}$}       & $X$       \\
    \midrule
    pMPC~\cite{yue2021pMPC}      &  \multicolumn{2}{c|}{$X_{-[t,t+k^\prime]}$ ($X_{[t,t+k^\prime]}$ is a phoneme)}       & $X_{[t,t+k^\prime]}$    \\
    \midrule 
    MPE~\cite{liu2020masked} & $X$ &  $Z_{-[t,t+k]}$  & $Z$    \\ %Lee: What is the difference between MPE and NPC??? And I believe it reconstruct the convolutional blocks output (learned target?)
    \midrule
    NPC~\cite{liu21l_interspeech}      & $X$  &   $Z_{-[t,t+k]}$  &    $X$   \\ %Lee: I am not 100% sure it is correct. Please check.
    \midrule
    \midrule
    \multicolumn{4}{c}{\textsc{Contrastive models}} \\
    \midrule
    \midrule
    Unspeech \cite{milde2018unspeech}       &   $X_{[t_1,t_2]}$ &   \textsc{-}   &  $X_{[t_0,t_1]}$,$X_{[t_2,t_3]}$ \\
    \midrule 
    CPC~\cite{oord2018representation}, wav2vec \cite{schneider2019wav2vec}, Modified CPC \cite{riviere2020unsupervised}         & $X_{[1,t]}$   &    \textsc{-}           & $z_{t+c},\, c\geq1$   \\ %Modified CPC is the same as CPC
    \midrule 
    Bidirectional CPC \cite{kawakami2020learning}      & $X_{[1,t]}$ or $X_{[t,T]}$ &  \textsc{-}    &    $z_{t+c}$ or $z_{t-c},\, c\geq1$   \\
    \midrule 
    vq-wav2vec \cite{Baevski2020vq-wav2vec}     &   $X_{[1,t]}$ &   \textsc{-}    &   $q_{t+c},\, c\geq1$   \\ 
    \midrule 
    wav2vec 2.0 \cite{baevski2020wav2vec}, wav2vec-C \cite{sadhu21_interspeech}\footnote{wav2vec-C adds reconstruction loss to wav2vec 2.0.}    & $X$             & $Z_{-[t,t+k]}$          & $Q_{[t,t+k]}$ \\
    \midrule 
    w2v-BERT \cite{w2vbert}     &$X$ &    $Z_{-[t,t+k]}$   &     $Q_{[t,t+k]}$ and $C_{[t,t+k]}$     \\
    \midrule
    Speech SimCLR \cite{SpeechSimCLR}\footnote{Speech SimCLR targets the latent representation of an augmented version of $X$ using a differently augmented $X$, and vice-versa.}    & \multicolumn{2}{c|}{$\textsc{augment}_1(X)$ and $\textsc{augment}_2(X)$}     &    $\textsc{augment}_2(Z)$ and $\textsc{augment}_1(Z)$   \\ 
    \midrule
    \midrule 
    \multicolumn{4}{c}{\textsc{Predictive models}} \\
    \midrule
    \midrule
    Discrete BERT~\cite{Baevski2020vq-wav2vec,baevski2019effectiveness} \footnote{Discrete BERT obtains codes $C$ from vq-wav2vec.}      &   \multicolumn{2}{c|}{$C_{-[t,t+k]}$}   & $C_{[t,t+k]}$  \\
    \midrule 
    HuBERT \cite{hsu2021hubert}\footnote{HuBERT is trained first using cluster IDs of the MFCCs as target and subsequently clusters IDs of the model representations from the last iteration.}, WavLM \cite{chen2021wavlm}\footnote{WavLM simulates noisy/overlapped speech as inputs.}  & $X$             & $Z_{-[t,t+k]}$          & $C_{[t,t+k]}$  \\ 
    \midrule
    data2vec \cite{data2vec}    & $X$             & $Z_{-[t,t+k]}$          & $\sum_{l}\bar{H}^{(l)}_{[t,t+k]}$  \\ 
    \midrule 
    BEST-RQ \cite{BEST-RQ}\footnote{BEST-RQ obtains codes $C$ by quantizing acoustic features using a random projection quantizer.}     &  \multicolumn{2}{c|}{$X_{-[t,t+k]}$}      &  $C_{[t,t+k]}$   \\ 
    \midrule 
    \bottomrule
\end{tabular}
}

% \end{minipage}
\label{table:pretext}
\end{table*}

\begin{comment}
\begin{table*}[ht!]
    \centering
    \caption{
    This table summarizes the approaches of the three categories of self-supervised learning. 
    Column (a) lists the names of the models and related references. 
    Column (b) defines the input to the models. 
    Column (c) defines any corruption of the input or some hidden representation. 
    Column (d) defines the target of the pretext task; the pretext task itself is described by the overall model category and the main text.
    $X=\{x_1,x_2,...,x_T\}$ is the input sequence in which $x_t$ can be an acoustic feature vector (e.g., MFCC, filterbank, or spectrogram features) or a waveform sample. 
    $X_{[t_1:t_2]}$ represents $\{x_{t_1},x_{t_1+1},...,x_{t_2}\}$.
    $X_{-[t_1:t_2]}$ represents $X$ with the segment $X_{[t_1:t_2]}=\{x_{t_1},x_{t_1+1},...,x_{t_2}\}$ masked.
    $x_{t}^{i}$ represents the $i$-th dimension of $x_t$.
    If $x_t$ is a frame in a spectrogram, then the $i$-th dimension corresponds to a specific frequency bin.
    $X^{-[f,f+j]}$ refers to a spectrogram $X$ but masked along the frequency axis from the $f$-th to $f+j$-th bin. 
    % $X_{-[t:t+k]}^{-[f,f+j]}$ refers to a spectrogram $X$ but masked along both time and frequency dimensions. 
    % $X^*$ is a temporally permuted version of $X$, that is, the $x_t$ are randomly shuffled to form $X^*$. 
    We indicate random temporal permutation of a sequence by indexing it with the set $\mathcal{P}_t\triangleq\textsc{permute}([0,t])$ where $\textsc{permute}(\cdot)$ returns a permutation of the given list. 
    We indicate data augmentation (e.g. reverberation) by the function $\textsc{augment}(\cdot)$. Subscripts indicate different augmentations. 
    % $X^\prime$ is an augmented version of $X$ (e.g., $X$ with reverberation), while $X^{\prime\prime}$ is $X$ adding distortion different from  $X^\prime$.
    $Z$ represents a localized latent representation sequence of $X$. %, and $Z^\prime$ and $Z^{\prime\prime}$ are the latent representation sequences of  $X^\prime$ and $X^{\prime\prime}$, respectively. 
    % $Z$ represents a localized latent representation sequence and $Z^\prime$ is the latent representation sequence of $X^\prime$. 
    $Z^{(l)}$ is $Z$ at the $l$-th layer of the model used to compute it.
    $\bar{H}$ is the contextualized sequence $H$ obtained from an exponential moving average (EMA) of the model undergoing training with no masking applied.
    $Q$ represents a sequence of quantized learned representations, and $C$ is a sequence of discrete cluster IDs.
    For contrastive models, we specify only positive targets.
    }
    \begin{minipage}{\textwidth}  
    \centering
    \renewcommand*\arraystretch{1.1}{
    \begin{tabular}{l|c|c|c}
    \toprule
    \textbf{Model} (a) & \textbf{Input} (b) & \textbf{Corruption} (c) & \textbf{Target} (d) \\
    \midrule
    \midrule
    \multicolumn{4}{c}{\textsc{Generative models}} \\
    \midrule
    \midrule
    Audio Word2vec~\cite{Chung2016AudioWord2Vec}, VQ-VAE \cite{vqvae}     & $X$ &   \textsc{-}    &  $X$  \\ %It has VQ as extra constraint.
    \midrule  
    Speech2Vec~\cite{chung2018speech2vec}, Audio2Vec~\cite{tagliasacchi2020pre} - skip-gram    & $X_{[t_1,t_2]}$  &     \textsc{-}   &    $X_{[t_0,t_1]}$,$X_{[t_2,t_3]}$     \\
    \midrule 
   Speech2Vec~\cite{chung2018speech2vec}, Audio2Vec~\cite{tagliasacchi2020pre} - cbow    & $X_{[t_0,t_1]}$,$X_{[t_2,t_3]}$   &     \textsc{-}   &    $X_{[t_1,t_2]}$   \\
    \midrule 
     PASE~\cite{pascual2019learning}, PASE+~\cite{ravanelli2020multi}\footnote{PASE uses multiple pretext tasks, but the authors find that reconstruction is most important.}       & $X$ &  \textsc{-}   &  Different modalities of $X$  \\
    \midrule 
    APC~\cite{chung2019unsupervised,chung20e_interspeech}         & $X_{[1,t]}$   & \textsc{-}              & $x_{t+c},\, c\geq1$    \\ %VQ-APC is the same as APC
    \midrule
    Speech-XLNet \cite{song20d_interspeech}     & \multicolumn{2}{c|}{$X_{\mathcal{P}_{t}}$}   &     $x_{i\sim\mathcal{P}^c_{t}}$  \\ %Lee: Use * to represent the permutation, hope it is not strange.
    \midrule  
    DeCoAR~\cite{ling2020deep}     & $X_{[1,t-1]}, X_{[t+k+1,T]}$ & \textsc{-} & $X_{[t,t+k]}$   \\
    \midrule
    Mockingjay~\cite{liu2020mockingjay}, Audio ALBERT~\cite{chi2020audio}, DeCoAR 2.0~\cite{ling2020decoar}   & \multicolumn{2}{c|}{$X_{-[t,t+k]}$}   & $X_{[t,t+k]}$    \\
    \midrule 
    TERA~\cite{liu2021tera}, BMR~\cite{wang2020unsupervised}  & \multicolumn{2}{c|}{$X_{-[t,t+k]}^{-[f,f+j]}$}       & $X$       \\
    \midrule
    pMPC~\cite{yue2021pMPC}      &  \multicolumn{2}{c|}{$X_{-[t,t+k^\prime]}$ ($X_{[t,t+k^\prime]}$ is a phoneme)}       & $X_{[t,t+k^\prime]}$    \\
    \midrule 
    MPE~\cite{liu2020masked} & $X$ &  $Z_{-[t,t+k]}$  & $Z$    \\ %Lee: What is the difference between MPE and NPC??? And I believe it reconstruct the convolutional blocks output (learned target?)
    \midrule
    NPC~\cite{liu21l_interspeech}      & $X$  &   $Z_{-[t,t+k]}$  &    $X$   \\ %Lee: I am not 100% sure it is correct. Please check.
    \midrule
    \midrule
    \multicolumn{4}{c}{\textsc{Contrastive models}} \\
    \midrule
    \midrule
    Unspeech \cite{milde2018unspeech}       &   $X_{[t_1,t_2]}$ &   \textsc{-}   &  $X_{[t_0,t_1]}$,$X_{[t_2,t_3]}$ \\
    \midrule 
    CPC~\cite{oord2018representation}, wav2vec \cite{schneider2019wav2vec}, Modified CPC \cite{riviere2020unsupervised}         & $X_{[1,t]}$   &    \textsc{-}           & $z_{t+c},\, c\geq1$   \\ %Modified CPC is the same as CPC
        \midrule 
    Bidirectional CPC \cite{kawakami2020learning}      & $X_{[1,t]}$ or $X_{[t,T]}$ &  \textsc{-}    &    $z_{t+c}$ or $z_{t-c},\, c\geq1$   \\
    \midrule 
    vq-wav2vec \cite{Baevski2020vq-wav2vec}     &   $X_{[1,t]}$ &   \textsc{-}    &   $q_{t+c},\, c\geq1$   \\ 
    \midrule 
    wav2vec 2.0 \cite{baevski2020wav2vec}, wav2vec-C \cite{sadhu21_interspeech}\footnote{wav2vec-C adds the reconstruction loss to wav2vec 2.0.}    & $X$             & $Z_{-[t,t+k]}$          & $Q_{[t,t+k]}$ \\
    \midrule 
    w2v-BERT \cite{w2vbert}     &$X$ &    $Z_{-[t,t+k]}$   &     $Q_{[t,t+k]}$ and $C_{[t,t+k]}$     \\
    \midrule
    Speech SimCLR \cite{SpeechSimCLR}\footnote{Speech SimCLR targets the latent representation of an augmented version of $X$ using a differently augmented $X$, and vice-versa.}    & \multicolumn{2}{c|}{$\textsc{augment}_1(X)$ and $\textsc{augment}_2(X)$}     &    $\textsc{augment}_2(Z)$ and $\textsc{augment}_1(Z)$   \\ 
   \midrule
   \midrule 
    \multicolumn{4}{c}{\textsc{Predictive models}} \\
    \midrule
    \midrule
    Discrete BERT~\cite{Baevski2020vq-wav2vec,baevski2019effectiveness} \footnote{Discrete BERT obtains codes $C$ from vq-wav2vec.}      &   \multicolumn{2}{c|}{$C_{-[t,t+k]}$}   & $C_{[t,t+k]}$  \\
    \midrule 
    HuBERT \cite{hsu2021hubert}\footnote{HuBERT is trained first using cluster IDs of MFCCs as target and subsequently cluster IDs of model representations from the last iteration.}, WavLM \cite{chen2021wavlm}\footnote{WavLM simulates noisy/overlapped speech as inputs.}  & $X$             & $Z_{-[t,t+k]}$          & $C_{[t,t+k]}$  \\ 
    \midrule
    data2vec \cite{data2vec}    & $X$             & $Z_{-[t,t+k]}$          & $\sum_{l}\bar{H}^{(l)}_{[t,t+k]}$  \\ 
        \midrule 
BEST-RQ \cite{BEST-RQ}\footnote{BEST-RQ obtains codes $C$ by quantizing acoustic features using a random projection quantizer.}     &  \multicolumn{2}{c|}{$X_{-[t,t+k]}$}      &  $C_{[t,t+k]}$   \\ 
        \midrule 
    \bottomrule
    \end{tabular}
    }

    \end{minipage}
    \label{table:pretext}
\end{table*}
\end{comment}

\begin{comment}
\begin{table*}[h!]
\centering

\setlength{\tabcolsep}{3pt}
\caption{
This table summarizes the generative approaches. 
Column (a) is the name of the models (if any) and their reference.
Column (b) is the corrupted speech signals for encoder input.
Column (c) is the target output of decoder.
Column (d) is regularization methods used if any.
Column (e) are the models using the same pretext in NLP or CV.
$S$ is a sequence of waveform samples, and $s_t$ is a single waveform sample.
$S^*$ is the degradation of $S$ with reverberation or additive noises. 
$S$ can be represented by a sequence of acoustic feature, $X=\{x_1,x_2,...,x_T\}$, in which $x_t$ is an acoustic feature vector like MFCC, fbank, spectrogram, etc. 
$X^*$ is the permuted version of $X$. That is, the frames in $X$ are randomly shuffled to form $X^*$. 
$X_{[t_1:t_2]}$ represents $\{x_{t_1},x_{t_1+1},...,x_{t_2}\}$.
$X_{-[t_1:t_2]}$ represents the segment $\{x_{t_1},x_{t_1+1},...,x_{t_2}\}$ in $X$ is masked, that is, the frames in the region is replaced by zero vectors or random sampled vectors.
$x_{t}^{i}$ represents the $i$-th dimension of frame $x_t$.
If $x_t$ is a frame in a spectrogram, then the $i$-th dimension corresponds to a specific frequency bin.
$X^{-[f_1,f_2]}$ means that for all the frames $x$ in $X$, their $f_1$-th to $f_2$-th dimensions are all masked. 
$X_{-[t_1:t_2]}^{-[f_1,f_2]}$ means $X$ are masked in two directions. 
}
\label{table:generative}
  {\renewcommand*\arraystretch{1.2}
\begin{tabular}{|l|l|l|l|l|}
\hline
(a) Model & (b) Corrupted Input & (c) Target Output & (d) Regularization &   (e) Counterparts \\ 
\hline \hline

APC~\cite{chung2019unsupervised, chung2020generative,chung2020improved} & $X_{[1,t-1]}$  & $x_{t+c},c\geq0$ & VQ~\cite{chung20e_interspeech} &  GPT~\cite{alex2018GPT,alex2019GPT2,brown2020gpt3} \\ \hline

%$X_{[1,t-1]}$  & $x_{t+c},c\geq0$ & VQ & VQ-APC~\cite{chung20e_interspeech} & -  \\

DeCoAR~\cite{ling2020deep} & $X_{[1,t-1]}$, $X_{[t+k+1,T]}$ &  $X_{[t,t+k]}$ & VQ~\cite{ling2020deep} &  ELMo~\cite{Matthew2018ELMO} \\
\hline

\tabincell{l}{ Mockingjay~\cite{liu2020mockingjay} \\ MPC~\cite{jiang2019improving,jiang2021further} \\ AALBERT~\cite{chi2021aalbert} \\
DeCoAR 2.0~\cite{ling2020decoar} } &
$X_{-[t,t+k]}$ & $X_{[t,t+k]}$ & VQ~\cite{ling2020decoar} &  \tabincell{l}{BERT ~\cite{jacob2019BERT}\\RoBERTa~\cite{Liu2019RoBERTa}} \\ \hline

pMPC~\cite{yue2021pMPC}   &
\tabincell{l}{$X_{-[t,t+k^\prime]}$\\($X_{[t,t+k^\prime]}$: phoneme-level segment)} & $X_{[t,t+k^\prime]}$ & - &  SpanBert~\cite{joshi-etal-2020-spanbert} \\ \hline

speech-XLNet~\cite{song20d_interspeech} &
$X^*_{[1:t-1]}$ & $x^*_t$ & - &  XLNet~\cite{Yang2019LNet}   \\ \hline

\tabincell{l}{NPC~\cite{liu21l_interspeech}\\MPE~\cite{liu2020masked}} &
\tabincell{l}{$X_{-[t,t+k]}$\\(mask on hidden layer)} & $X_{[t,t+k]}$ & - &   - \\ \hline

\tabincell{l}{BMR~\cite{wang2020unsupervised}\\TERA~\cite{liu2021tera}} &
$X_{-[t_1:t_2]}^{-[f_1,f_2]}$ & $X$ & dropout~\cite{luo2021drop} &  -  \\ \hline

\cite{chorowski2019unsupervised} &
$S$ & $S$ & \tabincell{l}{VQ-VAE~\cite{chorowski2019unsupervised}\\time-jitter regularizer~\cite{chorowski2019unsupervised}} & 
\tabincell{l}{\cite{van2017neural,VQVAE2} \\ (just name a few)} \\ \hline

\cite{tagliasacchi2019self, tagliasacchi2020pre} & 
$X_{[t-k,t-1]},X_{[t+k+1,t+2k]}$ & $X_{[t,t+k]}$  & - & CBoW~\cite{mikolov2013efficient}  \\ \hline

\cite{tagliasacchi2019self, tagliasacchi2020pre} &
$X_{[t,t+k]}$ & $X_{[t-k,t-1]},X_{[t+k+1,t+2k]}$ & - &  Skipgram~\cite{mikolov2013efficient} \\ \hline

\cite{quitry2019learning}  &
Magnitude of $X$ & Phase of $X$ & - &  - \\ \hline

TemporalGap~\cite{tagliasacchi2019self, tagliasacchi2020pre} &
$X_{[t_1,t_1+k]}$, $X_{[t_2,t_2+k]}$ & $|t_1 - t_2|$ & - &  -  \\ \hline

\cite{Carr2021Ranking} &
$X^*$ & correct order & - &  Jigsaw~\cite{noroozi2016unsupervised} \\ 
\hline

PASE~\cite{pascual2019learning} &
$S$ & $X$: LPS, MFCC, Prosody & - &  -  \\ \hline

PASE+~\cite{ravanelli2020multi} & 
$S^*$ & \tabincell{l}{$X$: LPS, MFCC, Prosody, \\ fbank, gammatone} & -  &  -  \\ \hline
\end{tabular}
}
\end{table*}
\end{comment}