### 2023-03-04 chapters/paper_review/stability.tex

%!TEX root = ../thesis.tex

\begin{table}[ht]
    \centering
    \resizebox{\linewidth}{!}{
    \begin{tabular}{cc|ccc}
        \toprule
        \multirow{2}{*}{feature} & \multirow{2}{*}{C} 
        & \multicolumn{3}{c}{PNMI (mean $\pm$ std) with K-means Training Size = } \\
        & & 1h & 10h & 100h \\
        \midrule\midrule
        \multirow{2}{*}{MFCC}
        & 100 & 0.251 $\pm$ 0.001 & 0.253 $\pm$ 0.001 & 0.253 $\pm$ 0.001 \\
        & 500 & 0.283 $\pm$ 0.001 & 0.285 $\pm$ 0.000 & 0.287 $\pm$ 0.001 \\
        \midrule
        \multirow{2}{*}{\textsc{Base}-it1-L6}
        & 100 & 0.563 $\pm$ 0.012 & 0.561 $\pm$ 0.012 & 0.575 $\pm$ 0.008 \\
        & 500 & 0.680 $\pm$ 0.005 & 0.684 $\pm$ 0.003 & 0.686 $\pm$ 0.004 \\
        \bottomrule
    \end{tabular}
    }
    \caption{Stability of K-means as an unsupervised unit discovery algorithm with respect to different features, numbers of clusters, and training data sizes. PNMI stands for phone-normalized mutual information.}
    \label{tab:stability}
\end{table}