### 2023-03-17 chapters/paper_review/05_zero.tex

%!TEX root = ../thesis.tex

\section{From Representation Learning to Zero Resources}
\label{sec:zero}

% {\color{black} Reviewer: TNS, Abdo, Hung-yi Lee }\\
% \kl{We discussed in our last meeting that I would add a pointer in this section to the acoustic word embedding section.  However, I'm not quite sure what the right spot for it is, so waiting for this section to stabilize a bit more first.}

In the SSL framework, speech representations can be
learned and used in various downstream tasks to achieve competitive, robust,
and transferable performance, as shown in
\crefrange{section:benchmark}{analysis}. 
However, labeled data is still required. 
For example, in ASR, utterances and their manual transcriptions are needed to learn downstream models or fine-tune representation models. 
Can a model learn without any labeled data? 
In \cref{secsec:unpaired}, we show how to learn ASR models without any paired audio and text and how SSL improves the framework.
In addition, many languages have no writing system. 
In \cref{subsec:zero}, the SSL representation is further used in scenarios where text data is unavailable.

\subsection{Unpaired text and audio} \label{secsec:unpaired}
% What are we going to do % how to match
%In the self-supervised learning framework, speech representations can be learned and used in the downstream tasks, but some labeled data for finetuning is needed.

\begin{table*}[ht]
\centering
    {\renewcommand*\arraystretch{1.4}}
\caption{
Unsupervised ASR.
TIMIT numbers are phoneme error rates (PER), while the numbers for
LibriSpeech are word error rates (WER).
SWC $=$ spoken word classifier, ST $=$ speech translation.
All speech and text are in English if not specified. 
\textcolor{black}{
The references in the table are sorted according to the date of publication.}
}
\label{fig:unsupervisedASR}
\resizebox{0.99\textwidth}{!}{
\begin{tabular}{cllllll}
    \toprule
Reference & Speech representation & Speech segmentation  & Text token/representation & Mapping approach & Refinement  & Results  \\ 
    \midrule
\cite{liu2018completely}& Audio word2vec~\cite{Wang2018SegAudioWord2Vec} & Oracle & Phoneme & \edit{Adversarial Training~\cite{gulrajani2017improved}} & - & TIMIT (PER): 63.6\%  \\ %Submitted on 1 Apr 2018
  \midrule
  \cite{chung2018unsupervised} & Speech2vec~\cite{chung_speech2vec_2018} & BES-GMM~\cite{kamper_segmental_2017} &  Word2Vec & \edit{Adversarial Training~\cite{conneau2017word}} & Self-training & SWC (Acc): 10.9\%  \\ %Submitted on 18 May 2018
%https://arxiv.org/pdf/1805.07467.pdf %The text embeddings were obtained by training Word2Vec on the transcriptions using the fastText implementation without subword information [3].
  \midrule

    \cite{chung2019towards} & \tabincell{l}{Speech2vec\\(English)} & Oracle & \tabincell{c}{Word2Vec\\(French)} & VecMap~\cite{artetxe-etal-2018-robust} & \tabincell{l}{LM rescore,\\sequence DAE} & ST (BLEU): 10.8\%   \\ %[Submitted on 4 Nov 2018]
\midrule

\cite{yeh2018unsupervised} & MFCC & GAS~\cite{wang2017gate} & Phoneme & Empirical-ODM~\cite{Liu2017ODM} & Self-training &  TIMIT (PER):  41.6\%  \\ %Submitted on 23 Dec 2018
  \midrule

\cite{chen2019completely} & MFCC & GAS  & Phoneme & \edit{Adversarial Training~\cite{gulrajani2017improved}} & Self-training &  TIMIT (PER):  33.1\%   \\
  \midrule

\cite{baevski2021unsupervised} & Wav2vec 2.0~\cite{baevski_wav2vec_2020} & \kmeans{} & Phoneme &\edit{Adversarial Training~\cite{gulrajani2017improved}} & Self-training & \tabincell{l}{TIMIT (PER):  18.6\%, \\ LibriSpeech (WER): 5.9\%}   \\
  \midrule

\textcolor{black}{\cite{Deciphering2022}} & \textcolor{black}{\tabincell{l}{Universal Phone \\ Recogniser}} & - & \textcolor{black}{Grapheme} & \textcolor{black}{Decipherment~\cite{ravi-knight-2011-deciphering}} & \textcolor{black}{Self-training} & \textcolor{black}{\tabincell{l}{GlobalPhone: 32.5\% to just 1.9\% \\ worse than supervised models} }  \\ %ubmitted on 12 Nov 2021 %m 32.5% to just 1.9% absolute worse than the equivalent fully supervised models trained on the same data
  \midrule
\textcolor{black}{\cite{w2v-u2}} & \textcolor{black}{Wav2vec 2.0~\cite{baevski_wav2vec_2020} }& - & \textcolor{black}{Phoneme} & \textcolor{black}{Adversarial Training~\cite{gulrajani2017improved}} & \textcolor{black}{Self-training} & \textcolor{black}{LibriSpeech (WER): 6.3\%}   \\
  \midrule

\textcolor{black}{\cite{w2v-u2}} & \textcolor{black}{Wav2vec 2.0~\cite{baevski_wav2vec_2020}} & - & \textcolor{black}{Grapheme} & \textcolor{black}{Adversarial Training~\cite{gulrajani2017improved}} & \textcolor{black}{Self-training} & \textcolor{black}{LJSpeech (WER): 64.0\%}  \\

%e MUSE [7] and VecMap [8], n
%The numbers in the section of unsupervised methods denoted as BLEU score (%) of VecMap /
%BLEU score (%) of MUSE 10.8 / 6.2
%11.3 / 7.3
%https://arxiv.org/pdf/1811.01307.pdf
    \bottomrule
\end{tabular}
}
\end{table*}

\subsubsection{Unsupervised ASR}
% {\color{black} Hung-yi}\\

If only unpaired speech and text are available, that is, the text is not a
manual transcription of speech, can the machine learn how to transcribe speech
into text?
This scenario is called \textit{unsupervised ASR}, and the framework is as
below. 
Given a set of unlabeled utterances $\mathcal{S}=\{S_1, S_2, ..., S_N\}$  and a
set of sentences $\mathcal{Y}=\{Y_1, Y_2, ..., Y_M\}$,\footnote{Note that
the speech and text are not paired, that is, $Y_i$ is not the transcription of
$S_i$.} a mapping function~$F$, which can take an utterance~$S$ as input and
generate its transcription, is learned from data. 
\Cref{fig:unsupervisedASR} summarizes recent work on unsupervised ASR,
including the speech representation used, the algorithm used to learn the
mapping without supervision, and the results. Below, we will discuss these
methods in more detail.

Adversarial training~\cite{goodfellow2014generative,arjovsky2017wasserstein, gulrajani2017improved} is one common way to learn such a 
mapping function. 
The framework includes a discriminator and a generator.
The mapping function~$F$ plays the role of the generator, which takes speech utterances as input and outputs text.
The discriminator learns to distinguish real text from the generated
output; the generator learns to ``fool'' the discriminator.
The generator and the discriminator are trained in an iterative, 
interleaved way. 
After the training, the generator serves as the speech recognition model.
\edit{
There is a large amount of work using gradient penalty in the objective of training discriminators~\cite{liu2018completely,chen2019completely,baevski2021unsupervised,w2v-u2}, which is inspired by Improved Wasserstein Generative Adversarial Network (WGAN)~\cite{gulrajani2017improved}.
}
\textcolor{black}{
Other ways to map speech and text include via segmental empirical output distribution matching (segmental empirical-ODM)~\cite{yeh2018unsupervised} and decipherment algorithm~\cite{Deciphering2022}.}

% Difficulty of unsupervised ASR (v.s. unsupervised MT)
Success in unsupervised neural machine translation
(MT)~\cite{artetxe2017unsupervised, conneau2017word, lample2017unsupervised}
has inspired innovative exploration of various unsupervised ASR algorithms.
If learning a translation model from unaligned sentences in two languages is
possible, considering speech and text as two different languages, learning the
mapping relationship from speech space to text space without an alignment 
% is not impossible.
  should likewise be possible.   % AMH: check
However, there are differences between unsupervised MT and unsupervised
ASR.
In unsupervised MT, most discrete source tokens can be mapped to specific
target tokens representing the same meaning. 
However, because speech has segmental structures, in unsupervised ASR, each text
token maps to a segment of consecutive acoustic features of variable length in
an utterance.
The generator is supposed to learn the segmental structure of an utterance
because information like token boundaries is not directly available.
This makes unsupervised ASR more challenging than unsupervised MT.

% how to represent audio
For unsupervised ASR to be feasible, the common idea is to make the speech and text units close to each other.
For the text side, word sequences can be transformed into phoneme sequences if a lexicon is available. 
On the other hand, we must first convert the speech signal into something close to phonemes. 
To achieve that, most studies on unsupervised ASR use a phoneme
segmentation module before the generator to segment utterances into
phoneme-level
segments~\cite{liu2018completely,chen2019completely,yeh2018unsupervised}. 
A representation vector or a token then represents each phoneme-level segment. 
It is easier for the generator to map each segment-level representation 
or token to the correct phoneme when the representation or token is highly correlated to the phonemes.
Wav2vec-U~\cite{baevski2021unsupervised} selects the input feature from different layers of wave2vec~2.0~\cite{baevski_wav2vec_2020}.
The selection criterion is based on analysis of the phonetic information in each layer.
\textcolor{black}{
If a universal phone recognizer trained from a diverse set of languages is available, it is another way to transcribe speech into phone-level tokens~\cite{Deciphering2022}.}
Another series of work is to transform a word into a word embedding. 
\cite{chung2018unsupervised,chung2019towards} use adversarial training to map the word-level
speech embedding space~\cite{chung_speech2vec_2018} to the word embedding space
and achieve promising performance on spoken word classification, speech
translation, and spoken word retrieval.
\Cref{fig:unsupervisedASR} summarizes the various ways to segment speech and represent speech and text in each reference.

%Hung-yi Lee: I comment this paragraph first.
\begin{comment}
%Hung-yi Lee: This paragraph is not finished yet.
\TNS{if we are limited on space,we can consider removing some of the detail of the various algorithms}
\TNS{split into 2 sentences, also have a connection to the previous para}
\cite{liu2018completely} is the first work to successfully achieve unsupervised
phoneme recognition by first clustering phoneme-level
embeddings~\cite{chung_audio_2016,Wang2018SegAudioWord2Vec} into a set of
tokens and then using GAN to learn the mapping relationship between tokens and
phonemes. 
However, in \cite{liu2018completely}, oracle phoneme segmentation is still
required to get reasonable performance. 
\cite{chen2019completely,yeh2018unsupervised} uses gate activation signals
(GAS)~\cite{wang2017gate} to get the phoneme-level segmentation, which is
unsupervised. 
%However, the loss term in~\cite{yeh2018unsupervised} includes an empirical average over all segments inside a logarithmic function, which can be biased if we sample this empirical average by a mini-batch average. \cite{liu2017unsupervised} therefore use an extremely large batch size during training to reduce the biasing problem.
%These problems are handled by another work~\cite{chen2019completely}.
%\cite{chen2019completely} removes the quantization step.
%Compared to~\cite{yeh2018unsupervised}, the discriminator in~\cite{chen2019completely} considers all possible information from the text dataset, not limited to the n-gram local statistics.
%Besides, \cite{chen2019completely} works well with a reasonable batch size, which makes it feasible when the computational resource is limited.
Wav2vec-U~\cite{baevski2021unsupervised} selects the input feature from
different layers of wave2vec 2.0~\cite{baevski_wav2vec_2020}, a self-supervised
model.
The selecting criterion is the PER by training a linear model in a supervised manner. 
%Should we mention this?
Wav2vec-U uses \kmeans{} to cluster the selected features, and the boundaries
are drawn whenever the clustered index changes.
%Wave2vec-U tends to generate more segments compared to the oracle phoneme-level segments. For example, their segmentation gives 0.93 precision, the ratio of the oracle phoneme-level boundaries being hit, and 0.38 recall, the ratio of the generated boundaries hit the oracle phoneme-level boundaries.
While Wav2vec-U merges the neighboring segments containing the same predicted
labels in each step of GAN training, this design can be viewed as refining the
segmentation implicitly.  
%Should we mention this?
%At generator output, the wav2vec-U will combine the neighboring predicted posterior containing the same predicted label into a predicted posterior.
\end{comment}

%In \cite{chung2018unsupervised, chung2019towards}, the word boundaries are also generated automatically with speech only. 
%However, \cite{chung2018unsupervised, chung2019towards} need oracle boundaries, therefore not completely unsupervised as the framework proposed in this paper.

% more improvement: iterative training, development 
As shown in \cref{fig:unsupervisedASR}, most studies use
\textit{self-training} to refine the models.
% to achieve better performance.  % AMH: redundant
In self-training, the generator serves as the first-version phoneme recognition model.
Inputting unpaired speech to the generator generates the corresponding
``pseudo transcription''. 
We then view the speech utterances and their pseudo transcriptions as paired
data which we use to train a model in a supervised manner.
Although the pseudo transcriptions have more errors than oracle
transcriptions, experiments show that training models on pseudo
transcriptions still significantly boosts performance compared to the
first-version model.
%With the new model, we can further obtain new transcriptions.
%The iteration will continue until the performance converges.

%Using performance to end this section.
Wav2vec-U~\cite{baevski2021unsupervised} achieved state-of-the-art results at the time, which suggests that representation learning is
essential for the success of unsupervised ASR.
It achieved an 11.3\% phoneme error rate on the TIMIT benchmark. 
On the LS benchmark, wav2vec-U achieved a 5.9\% \edit{WER} on
\emph{test-other}, rivaling some of the best published systems trained on 960 hours of labeled data from only two years earlier. 
\textcolor{black}{
And wav2vec-U 2.0~\cite{w2v-u2} further removes the requirement of the segmentation stage, so the unsupervised ASR model can be learned in an end-to-end style.}
The robustness of wav2vec-U was further analyzed with respect to 
domain-mismatch scenarios in which the domains of unpaired
speech and text were different~\cite{lin2021analyzing}.
Experimental results showed that domain mismatch leads to inferior performance, but a representation model pre-trained on the targeted speech domain extracts better representations and reduces this drop in performance.

%Hung-yi: The technique below is very important. But I think it is fine to ignore this part in a reviewe paper due to space limitation. 
\begin{comment}
In unsupervised learning, how to determine the hyperparameters is an issue. 
In a typical supervised framework, hyperparameter tuning is done by a
development set, which includes labeled data.
In an unsupervised setting, the existence of this kind of development is tricky. %Is tricky a correct term?
The existence of such kind of development set means the presence of the labeled
data.
Therefore, wav2vec-U~\cite{baevski2021unsupervised} develops a new approach to
determine the hyperparameters without a labeled development set. 
%should I provide more details? 
\end{comment}

%Hung-yi: A series of work (most from MS) focuses on meager resources but not wholly unsupervised. Perhaps we have to mention the papers.

\subsubsection{ASR-TTS}
%{\color{black} Shinji}\\
%{\color{black} Hung-yi: In the last meeting, we discussed the connection between this section and self-supervised learning. In addition to looking for papers using self-supervision to improve the ASR-TTS framework here, I think there may be another direction. Can we regard this ASR-TTS framework as a kind of self-supervised learning? ASR-TTS is similar to an autoencoder, and the output of ASR is "latent representation". In other words, this section describes a special self-supervised learning method, whose representation is expressed as "text."}\\
%{\color{black} Hung-yi: However, the above idea sounds tricky. In addition, the ``ASR-TTS'' autoencoder is learned by semi-supervised, not self-supervised.}
% AMH: to make the paper more consistent, I recommend changing all $f^{\mathrm{asr}}$ to $f_{\mathrm{asr}}$
Here we describe an alternative approach by which to train an ASR and \edit{text-to-speech (TTS)} system
based on unpaired text and audio. The ASR-TTS framework, which \edit{combines the ASR} and
TTS systems in a cascaded manner,
can be regarded as an autoencoder, where the encoder~$f$
corresponds to the ASR module and the decoder~$g$ corresponds to the TTS module.
In this framework, we consider the intermediate ASR output as a latent
representation; the framework as a whole can be regarded as a variant of
self-supervised learning.\footnote{However, to make this complicated
system work, we often require that data is paired. Therefore, in practice,
ASR-TTS and other methods described in this section are categorized as
semi-supervised learning.}

The ASR-TTS framework can jointly optimize both ASR and TTS \edit{without using paired data}~\cite{tjandra2017listening,hori2019cycle,wang2020improving}.
A speech chain~\cite{tjandra2017listening,tjandra2018machine} is one 
successful way to utilize audio-only and text-only data to train both
end-to-end ASR/TTS models.
This approach first prepares pre-trained ASR model \edit{
$f_{\mathrm{asr}}(X)$ with acoustic input $X$ and pre-trained TTS model $g_{\mathrm{tts}}(Y)$ with text input $Y$.
By following the TTS system with an ASR system, we generate new 
acoustic feature sequence~$\hat{X}$, which must be close to the original input~$X$.
Thus, we design a loss function $\mathcal{L}_{\mathrm{asr} \rightarrow \mathrm{tts}}(X, \hat{X})$, where $\hat{X}$ is generated by
%and~$\hat{X}$:
%\begin{equation}
%    \mathcal{L}_{\mathrm{asr} \rightarrow \mathrm{tts}} (X, \hat{X}),
%\end{equation}    
%where
\begin{equation}
    \hat{X} = g_{\mathrm{tts}}(f_{\mathrm{asr}}(X)).
    \label{eq:asr-tts}
\end{equation}
Thus, we train the ASR model (or both ASR and TTS models) using only
the acoustic input by minimizing $\mathcal{L}_{\mathrm{asr} \rightarrow \mathrm{tts}}$.  
%\begin{equation}
%\hat{\theta}_{\mathrm{asr}} = \arg \min _{\theta_{\mathrm{asr}}} \mathcal{L}(Y, \hat{Y}; \theta_{\mathrm{asr}}, \theta_{\mathrm{tts}}).
%\end{equation}
Note that this approach does not require the supervised text data~$Y$}.
As an analogy to the generative approach in \cref{sec:generative}, the
intermediate ASR output~$\hat{Y}$ can be regarded as the latent representation~$Z$.

The other cycle with the text-only data~$Y$ is also accomplished by the
concatenated TTS-ASR systems:\edit{
\begin{equation}
    \hat{Y} = f_{\mathrm{asr}}(g_{\mathrm{tts}}(Y)).
    \label{eq:tts-asr}
\end{equation}
Similarly, this approach does not require the supervised audio data~$X$, and the
intermediate TTS output~$\hat{X}$ can be regarded as the latent representation~$Z$}.
Although this approach initially freezes either the ASR or TTS model, 
extensions of this study~\cite{hori2019cycle,tjandra2019end,baskar2019semi}
implement the joint training of both ASR and TTS parameters using 
REINFORCE~\cite{williams1992simple} and straight-through estimators.

An emerging technique uses a well-trained TTS system to generate
speech and text data from text-only data.
This technique is a sub-problem of the TTS-ASR approach formulated in
\eqref{eq:tts-asr} in which we fix the TTS system part and estimate only the ASR
parameters. 
For example, a huge amount of text resources can be obtained from the web
and document archives without corresponding audio data.
%Self-supervised learning in this paper mostly focuses on representation learning perspectives of such unpaired data.
The typical use case scenario of such a text resource for ASR is through \edit{the
language model}. 
We combine the ASR and language model via a noisy channel 
model~\cite{jelinek1997statistical}, a weighted finite state 
transducer~\cite{mohri2002weighted}, or shallow 
fusion~\cite{gulcehre2015using,chorowski2016towards}.
However, the progress of TTS systems boosted by deep 
learning~\cite{oord_wavenet_2016,shen2018natural} has inspired another interesting and
straightforward research direction: \edit{artificially creating paired text and audio
data $\{\hat{X}, Y\}$ with only text data~$Y$ by generating the corresponding
audio data~$\hat{X}$ with TTS.} %\begin{equation}
%    \{\hat{X}, Y\} \text{ where } \hat{X} = g_{\mathrm{tts}} (Y).
%\end{equation}
The most straightforward approach is to simply use multi-speaker TTS to
generate the waveform with various acoustic 
variations~\cite{li2018training,ueno2019multi,rosenberg2019speech,laptev2020you,huang2020using}.
The other approaches are based on the generation of high-level (more
linguistic) features instead of generating the waveform, e.g., encoder
features~\cite{hayashi2018back} and phoneme 
features~\cite{renduchintala18_interspeech,masumura20_interspeech}.
This approach is similar to the back-translation technique developed in \edit{neural machine translation}~\cite{sennrich2016improving}.
One benefit of the above data generation approaches is that it can be used to feed
unseen word or context phrases to end-to-end ASR.

% {\color{black} Shinji: puts some notions to encourage the community to contributions to work on this direction and has more connections to the current SSL.}

\subsection{No text or lexicon} \label{subsec:zero}

% \kl{moved the acoustic word embedding subsection from here to the rep learning methods section.}\\

\subsubsection{Zero-resource speech technologies and challenges}
% {\color{black} Shinji}\\
\label{zero_speech}
Zero-resource speech technologies, which seek to discover linguistic concepts
from audio only (no text nor lexicon), are one of the most active applications
of unsupervised/self-supervised speech processing.
Zero-resource speech technologies were initially studied for acoustic and
linguistic unit discovery from speech data without linguistic resources,
e.g., transcriptions and other annotations~\cite{jansen2013summary}.
This study was motivated by unsupervised query-by-example, applications of
non-parametric Bayesian machine learning to speech processing, and low-resource
speech recognition, and was also inspired by the learning process of infants.
The goal of this type of work is to build spoken dialog systems in a zero-resource
setup for any language.
% To facilitate the zero-resource problem in the community, 
  To encourage      zero-resource research,                  % AMH: check
zero-resource speech challenges have been organized since 2015.

In this section, we describe the research directions of zero-resource speech
technologies by following the series of zero-resource speech challenges.
\begin{itemize}
	 \item \edit{Zero Resource Speech Challenge 2015~\cite{versteegh2015zero} mainly focused on building an acoustic model without using any
	 linguistic annotations based on subword unit modeling and spoken term
	 discovery tracks}.
	 \edit{For the subword unit modeling track, the ABX score for the within- and across-speaker
	 tasks was used as an evaluation metric.}
	 The spoken term discovery track used the normalized edit distance and coverage
	 scores in addition to the precision, recall, and F1 scores for types,
	 tokens, and boundaries.
    Both tracks were based on the English and Xitsonga languages.
	 \item The Zero Resource Speech Challenge 2017~\cite{dunbar_zero_2017} focused on 
	 unseen language and speaker aspects from the previous challenge. For
	 example, to demonstrate the robustness against unseen languages, the systems
	 were developed with English, French, and Mandarin and tested on
	 two ``surprise'' languages: German and Wolof.
	 Similarly, robustness against unseen speakers was demonstrated by varying
	 the amount of speech available for each speaker.
	 \item The Zero Resource Speech Challenge 2019~\cite{dunbar_zero_2019} extended a goal of
	 previous challenges by synthesizing speech without
	 text or phonetic labels but with acoustic units obtained using
	 zero-resource techniques.
	 The evaluation metrics were also extended to subjectively evaluate the
	 quality of synthesized speech, including its intelligibility, naturalness,
	 and speaker similarity.
	 \item The Zero Resource Speech Challenge 2020~\cite{dunbar_zero_2020} was based on two
	 tracks, revisiting previous challenges with different evaluation metrics. 
	 The first task revisited the 2019 challenge with low bit-rate subword
	 representations that optimize the quality of speech synthesis. The second
	 task revisited the 2017 challenge by focusing on the discovery of word-like
	 units from unsegmented raw speech.
	 \item The Zero Resource Speech Challenge 2021~\cite{nguyen2020zero}, the latest
	 \edit{challenge}, expanded the scope to include language modeling tasks.
	 In addition to phoneme-level ABX, the challenge includes lexical,
	 semantic, and syntactic evaluation metrics computed via a language model of
	 pseudo-acoustic labels.
\end{itemize}
These challenges have facilitated the tracking of technical trends in
zero-resource speech technologies.
For example, research directions thereof 
have expanded to various speech processing components to cover the entire
spoken dialogue systems.
To keep up with this expansion, the challenge has continued to develop
appropriate evaluation metrics for zero-resource scenarios.
Following the success of representation learning, baseline and challenge
techniques have shifted from purely generative 
models~\cite{ondel2016variational,heck2017feature}, deep 
autoencoders~\cite{tjandra19_interspeech,chorowski_unsupervised_2019}, and incorporation of
neural-network-based TTS/VC techniques~\cite{vanniekerk_vectorquantized_2020} to
self-supervised learning~\cite{maekaku21_interspeech}.
The latest challenge included the visual modality, continuing the
expansion to include more aspects of human interaction.

\subsubsection{Textless NLP}
% {\color{black} Abdo}\\

Textless NLP is a new research direction that leverages the progress mentioned
above in self-supervised speech representation learning to model language
directly from audio, bypassing the need for text or 
labels~\cite{nguyen_generative_2022,polyak_speech_2021,kharitonov_textfree_2021,emotion_conversion}. 
Not only does this open the gate for language and dialect modeling without
orthographic rules, but it also offers the opportunity to model other
non-lexical information about how speech is delivered, e.g., speaker identity,
emotion, hesitation, interruptions. 
The generative spoken language model (nguyen_generative_2022)~\cite{nguyen_generative_2022} utilizes discrete
representations from wav2vec~2.0, HuBERT, and CPC algorithms as inputs to an
autoregressive language model trained by \edit{using the cross-entropy function} to maximize the
probability of predicting the next discrete speech token. A \edit{synthesis} module
follows the language model to produce speech waveforms given the generated
discrete speech units. The generated spoken continuations compete with
supervised generations and synthesis using a character language model in
subjective human evaluations. The model completes incomplete words
(pow[..] $\rightarrow$ POWER) and continues using words in the same general mood (dark $\rightarrow$
BLACKNESS)\footnote{https://speechbot.github.io/gslm/} and has been extended to
model and generate
dialogues~\cite{dialogue_nguyen_generative_2022}.\footnote{https://speechbot.github.io/dgslm/}
%\cite{}. 
Given its flexibility in modeling spoken content, the nguyen_generative_2022 has been further extended
to jointly model content and prosody~\cite{kharitonov_textfree_2021}. This prosodic-nguyen_generative_2022 model
introduced a multistream causal Transformer, where the input and output layers
use multiple heads to model three channels:  discrete speech units, duration,
and quantized pitch. The prosodic-nguyen_generative_2022 model jointly generates novel content
and prosody congruently in the expressive 
style of the prompt.\footnote{https://speechbot.github.io/pgslm/}
%\cite{}. 
Going one step further, \cite{emotion_conversion} used a speech emotion
conversion framework to modify the perceived emotion of a speech utterance
while preserving its lexical content and speaker identity. Other studies have
extended the idea of textless language processing or audio discrete representation to applications such as 
spoken question answering~\cite{spoken_qa_dual}, speech separation~\cite{shi2021discretization}, TTS~\cite{hayashi2020discretalk}, and speech-to-speech
translation~\cite{textless_translation}.