### 2023-08-09 chapters/paper_automated.tex
Finalize model-agnostic paper and add stroke paper

%!TEX root = ../thesis.tex

\chapter[automated medical coding on mimic-iii and mimic-iv: a critical review and replicability study]{Automated Medical Coding on MIMIC-III and MIMIC-IV: A Critical Review and Replicability Study}
\label{chp:paper-retrospective}
\ifthenelse{\equal{\skippapers}{true}}{}{

\section*{Abstract}

}