### 2023-08-26 appendices/paper_brief.tex
Some formatting and changes to labels

%!TEX root = ../thesis.tex

\chapter[a brief overview of unsupervised neural speech representation learning]{A Brief Overview of Unsupervised Neural Speech Representation Learning}
\label{app:paper-brief}
\ifthenelse{\equal{\skipappendices}{true}}{}{

\todo[inline]{Include brief overview of unsupervised neural speech representation learning in appendix.}

}