### 2023-10-02 frontmatter/acknowledgements.tex

%!TEX root = ../thesis.tex
\chapter[acknowledgements]{Acknowledgements}
First of all, I want to thank all my supervisors... and all my friends too.

% Supervisors: Jes, Lars

% DTU Group:

% Corti: 
% - Lars, Andreas, Michael for letting me in.
% - Machine learning team for supporting me in my first job and to take on the challenge of a PhD project.

% - Joakim, Lasse,

% family, friends and Rikke: support, wonderful being and unwavering and tolerant nature she has supported me in this undertaking at every step of the way.

\vspace*{\fill}
\noindent \textit{The project was funded by Corti and Innovation Fund Denmark through industrial PhD grant no.\@ 0153-00167B.}