### 2023-10-15 chapters/discussion.tex
Update discussion and background

%!TEX root = ../thesis.tex

\chapter[discussion]{Discussion}\label{chp:discussion}
% ~5 pages
%
% OUTLINE:
% - paper-hierarchical:
%   - Suitability of VAEs for representation learning (minimization of mutual information and sensitivity to implicit prior such as architecture)
% - paper-benchmarking
%   - Inferiority of probabilistic methods compared to self"=supervised learning. 
% -

The rapid progress of machine learning that started about a decade ago with the 2012 ImageNet competition and the work of \textcite{krizhevsky_imagenet_2012} all but slowed down during the course of this project. 
Following the papers by \textcite{choi_waic_2019,nalisnick_detecting_2019,hendrycks_deep_2019}, the field of out-of-distribution detection saw a surge of interest that has since grown yearly, while self"=supervised learning for speech has continued to advance with works such as CPC \parencite{oord_representation_2018}, wav2vec 2.0 \parencite{baevski_wav2vec_2020} and data2vec \parencite{baevski_data2vec_2022}.

Having focused on VAEs for out-of-distribution detection in \cref{part:unsupervised-uncertainty-estimation} and on self"=supervised methods for representation learning in \cref{part:self-supervised-speech-representation-learning}, this discussion will bridge a gap between the two approaches and consider how VAEs might be enabled to learn representations more competitive with self"=supervised methods. 
Furthermore, since the work in \cref{part:medical-applications} only had limited focus on uncertainty, we will also present and discuss the use of calibration techniques for the stroke recognition model drawing connections to how such systems are perceived and used in practice.

\section{Representation learning with variational autoencoders}
%
In this thesis we studied two different approaches to speech representation learning: VAEs in \cref{chp:paper-hierarchical,chp:paper-modelagnostic,chp:paper-benchmarking} and self"=supervised methods in \cref{chp:paper-review}. 
As we saw, the probabilistic formulation of VAEs provides benefits for their application to uncertainty quantification, although competitive methods that use self"=supervised foundation models have also been proposed \textcite{xiao_we_2021,hendrycks_using_2019,bergman_classificationbased_2020}. Self"=supervised methods are generally superior when it comes to performance on downstream tasks. While this of course depends on the task and the amount and type of unlabeled and labeled data, self"=supervised methods for speech are commonly evaluated on downstream tasks that are harder and more complex than those used for VAEs. For instance, between \cref{tab: phoneme recognition (PER),tab:unsupervisedASR} self"=supervised methods can be seen to outperform VAEs for phoneme recognition while self"=supervised methods are also reported for word-level speech recognition. 
In the following, we will discuss potential directions of future research that might help shed light on why VAEs representations underperform on downstream tasks and how they might be improved.

% Considering VAEs and self"=supervised methods, it seems we have not yet been able to build a model that has a principled probabilistic representation of uncertainty without compromising its ability to learn representations that are useful for downstream applications.

% Different from self"=supervised methods, VAEs learn the distribution over the training data and are forced to encode all aspects important to it. 
% As we saw in \cref{chp:paper-hierarchical}, this will generally include low-level features that are important for accurate input reconstruction, but usually of lesser interest for downstream tasks.

\paragraph{Designing the latent space} 
Architectural improvements and hierarchies of latent variables are two related, and interesting, approaches to learning more semantic features in VAEs. 
Some success has been demonstrated in the image domain with models like BIVA \parencite{maaloe_biva_2019}, NVAE \parencite{vahdat_nvae_2020}, and VD-VAE \parencite{child_very_2021}, with up to 78 stochastic layers, achieving tight likelihoods and high-quality samples, although limited effort has put towards using such models for downstream tasks. 
For speech, a hierarchical model operating on multiple temporal scales, such as the Clockwork VAE \parencite{saxena_clockwork_2021} adapted for speech in \cref{chp:paper-benchmarking}, might help better capture dependencies at longer ranges and encode more semantic features. 
For instance, phonetic content for pronunciation might be learned at lower layers, speaker identity at the upper layers, and semantic features, such as word-meaning, in between. 
Although some work has successfully separated speaker identity from content by modifying the latent variables and their dependencies \parencite{hsu_unsupervised_2017}, models that can learn a deep hierarchy of features for speech remains an open challenge. The work presented in \cref{chp:paper-benchmarking} represents an effort to make progress towards such a model.

\paragraph{Adding a few labels} 
Another way to improve the representations learned in VAEs is via semi-supervised learning. 
Here, a few labels are used to inform which patterns that are learned from a large, mostly unlabeled, data set. 
This is usually done by defining a new stochastic variable as the target and deriving a semi-supervised version of the ELBO that accommodates using the labels when they are available, or marginalizing the target variable when they are not. The VAE is then trained on the labeled and unlabeled data simultaneously.

Although VAEs are strong models for semi-supervised learning \parencite{kingma_semi-supervised_2014,kingma_improved_2016,maaloe_biva_2019}, self"=supervised methods have established themselves as superior for most tasks in this setting \cite{baevski_wav2vec_2020,jiang_speech_2021, liu_learning_2023}. 
Despite being theoretically appealing, joint objectives for semi-supervised learning in VAEs have practical drawbacks. By training on labeled and unlabeled data simultaneously, semi-supervised VAEs are often less flexible than are self"=supervised methods that divide the training into two separate phases. By first fitting a general foundation model in an expensive pre"=training phase it can then later be fine"=tuned for many downstream tasks, at relatively low cost. 
VAEs, on the other hand, must often learn the labeled downstream tasks while simultaneously training on the unlabeled data. This is computationally expensive, but also requires retraining with the unlabeled data, when new supervised data becomes available or a new task is added.

\paragraph*{Approximating less}
VAE are inherently approximate. Training is performed on a lower bound of the likelihood and inference is variational, amortized and for non-hierarchical models, mean field. As a consequence, the gradient itself is estimated inexactly, usually by only one posterior sample, leading to non-negligible variance. 
For that reason, it seems plausible that using tighter bounds and reducing gradient variance improve representation learning in VAEs. 
Importance weighting the ELBO \parencite{burda_importance_2016} provides a tight bound on the likelihood, reduces gradient variance, and induces a complex implicit posterior distribution. 
Its use has become standard when reporting likelihood benchmarks but as we described in \cref{sec:variational-autoencoders}, \textcite{rainforth_tighter_2019} demonstrated that using it during training introduces high gradient variance for the inference network hurting its ability to learn useful representations. Later works largely solved this issue \parencite{roeder_sticking_2017,tucker_doubly_2019,bauer_generalized_2021} and showed that the reduced variance leads to improved likelihoods.

Despite much of this progress being prior to the latest, very large hierarchical VAEs, such as BIVA \parencite{maaloe_biva_2019}, NVAE \parencite{vahdat_nvae_2020}, and VD-VAE \parencite{child_very_2021}, these models are all trained with the single-sample ELBO using regularly reparameterized gradient estimator. Instead, these models rely on advanced inference networks \parencite{maaloe_biva_2019} and architectural improvements \parencite{vahdat_nvae_2020, child_very_2021} to overcome the challenges of training hierarchical VAEs. 
This seems to indicate that there might be untapped potential in consolidating the work on low-variance gradient estimators and advanced architectures and inference networks.

% Although the path-wise derivative has lower variance than most alternatives, such as score function estimators, a number of works have attempted to reduce its variance.
% \textcite{roeder_sticking_2017} note that the path-wise derivative can be factored into a path-wise estimator and a score function estimator and propose to ignore the score function factor to reduce gradient variance. 
% \textcite{rainforth_tighter_2019} show that this problem is even larger for the importance weighted ELBO \parencite{burda_importance_2016} and leads to diminishing returns when increasing the number of samples. 
% Rather than removing the score function factor in the estimate, \textcite{tucker_doubly_2019} propose to reparameterize it and derive low variance gradient estimators, also for importance weighted ELBO.

\paragraph{Masking} 
As discussed in \cref{chp:paper-review}, masking is one of the driving techniques behind the success of self"=supervised methods for speech \parencite{devlin_bert_2018,baevski_wav2vec_2020}. By removing tokens from the input or an early feature extraction layer and tasking a model with inferring their representations, models are forced to learn how neighboring tokens relate to those masked. Depending on the size of the mask, these dependencies can be more or less semantic in nature. For instance, the wav2vec 2.0 model is well-known to learn representations that have high similarity with word identity and meaning \parencite{pasad_layerwise_2021}. 
In comparison, reconstruction in VAEs is done from latent variables that are inferred from the full, unmasked input. Since all information is generally available, this might allow the encoder and decoder models to perform well for reconstruction, even with no, or limited, use of contextual dependencies. Furthermore, VAE reconstruction almost always targets the direct input in order to learn the distribution over the training data and to enable generating new samples. However, compared to many self"=supervised methods that target intermediate, learned representations, this forces VAEs to encode all aspects of the input that are important to its distribution. As we saw in \cref{chp:paper-hierarchical}, this will generally include low-level features that are necessary for accurate input reconstruction. Such features require large latent space representations and model capacity \parencite{vahdat_nvae_2020,child_very_2021}, but are usually of lesser interest for downstream tasks \parencite{baevski_wav2vec_2020}.

Since the ELBO does not immediately allow for masking as part of the training, it has been only sparsely examined for VAEs. Specifically, masking has seen the most attention for VAEs within missing data imputation. In this setting, the input is partially observed, and often represented as a segmentation into observed and missing parts via a mask that indicates where the data is missing. The model is then trained to infer the latent variable from the observed data and reconstruction also deals only with the observed data. 
By comparison to self"=supervised approaches that use masking for representation learning, the missing data setting of VAEs focuses on reconstructing the observed data rather than the missing, which might not lead to the same benefits in representation learning. 
The idea of using VAEs to impute missing data was already examined in the seminal paper by \textcite{rezende_stochastic_2014}. Here the model was trained with fully observed data and used to impute data in an iterative sampling approach post hoc, leaving the learned representations unchanged.
Previous work that trains on partially observed data has largely focused on the ability of these models to yield high-quality imputations within the tabular and image data domains and have not probed for the effects on the learned latent representation \parencite{mattei_miwae_2019, ipsen_not-miwae_2021}.

Including masked objectives into the principled probabilistic framework of VAEs might require creative thinking, but it seems likely that relaxing the requirement of exact input reconstruction and enforcing the need of encoding (and decoding) contextual information could lead the way for future types of VAEs to learn representations that are competitive with self"=supervised methods.

% \subsection{Uncertainty estimation with self-supervised methods}
% %

% \textcite{pasad_layerwise_2021} found that layers 6-7 hold the most information about phonetic content and word identity and meaning for \texttt{wav2vec2-base}. 
% For \texttt{wav2vec2-large} the phonetic content is highest in layers 11 and 18-19 with a drop in between, while word identity and meaning are highest between layers 12 and 18.

% Uncertainty of representations versus uncertainty of downstream task outputs. 
% % \textcite{wickstrom_relax_2023} propose a method for quantifying uncertainty based on

% \section{\Cref{chp:paper-hierarchical} revisited: \dots}

% \lesstodo[inline]{Discuss sensitivity to ``implicit" prior such as architecture (and probably optimization method and other). Include reference to \parencite{huszar_is_2017} discussing the usefulness of using a maximum likelihood objective for representation learning in generative models.}
% \lesstodo[inline]{Discuss whether VAEs are even suitable for representation learning due to them minimizing a mutual information term in the ELBO (derive this form).}
% \lesstodo[inline]{Discuss some recent work e.g. \cref{morningstar_density_2021}}

% \section{\Cref{chp:paper-benchmarking} revisited: Bested?}

% \lesstodo[inline]{Discuss whether variational autoencoders are viable for learning good representations for speech for downstream tasks when alternatives such as SSL method exist.}
% \lesstodo[inline]{Out-of-distribution detection on speech?}

% \section{\Cref{chp:paper-review} revisited: \dots} \label{sec:discussion-paper-review}

% \lesstodo[inline]{Are self"=supervised speech representations useful for unsupervised uncertainty estimation? \parencite{nava_stateconsistency_2021, nava_uncertaintyaware_2021} Within robotics but not really related.}
% % Since the main focus within self"=supervised learning has been on improving downstream task performance, very limited work, if any, has investigated self"=supervised representations in terms of uncertainty estimation. 
% % However, in the context of medical applications where data can be abundant but labels are sparse, unsupervised uncertainty estimation is a very interesting direction for future work.
% \lesstodo[inline]{OOD data: Generalization or detection (https://arxiv.org/pdf/2110.11334.pdf)?}

\section{\Cref{chp:paper-retrospective} revisited: Calibration}

\begin{figure}
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_curve_ensemble_and_all_models_uncalibrated.pdf}
        % \caption{}
        % \label{fig_discussion:calibration_curve_ensemble_and_all_models_uncalibrated}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/histogram_ensemble_and_single_model.pdf}
        % \caption{}
        % \label{fig_discussion:histogram_ensemble_and_single_model}
    \end{subfigure}
    \caption[Calibration curve for the uncalibrated stroke recognition ensemble and empirical distribution of predicted probabilities.]{ Calibration curve for the uncalibrated stroke recognition ensemble (left) and the histogram of predicted probabilities (right) for the test set. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model}
\end{figure}

Our work on stroke recognition in \cref{chp:paper-retrospective} focuses on the predictive performance of the ensemble model and an analysis of feature importance, but does not explicitly consider uncertainty estimation. 
As we discussed in \cref{subsec:model-calibration}, such a model can be calibrated to predict probabilities that are aligned with the empirical probability of the model being correct on some validation set. Here we present and discuss model calibration for the ensemble model.

We compute the calibration curve by sorting the probabilities predicted on the test set into a number of bins spanning the range from zero to one. For each bin $b$, we compute the mean predicted probability $\bar{p}$ and the fraction of examples for which the model predicted correctly $r_{b}$. The calibration curve is the drawn from the $\{(\bar{p}_{b}, r_{b})\}_b$ pairs. For any given bin, a perfectly calibrated model would have the same fraction of correct predictions as that bin's mean value, $\bar{p}_{b} = r_{b}\,\forall\,b$.

The calibration curve for the uncalibrated stroke recognition ensemble and its constituent models is plotted in \cref{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model} along with a histogram of its predicted probabilities. The miscalibration issue that we previously discussed is clearly visible as a strong overconfidence for both ensemble and constituents, although the ensemble is much better calibrated than its constituents. Since the ensemble's output probability is computed as the harmonic mean of the five constituent model probabilities, it can never exceed the maximum probability predicted between the constituent models. This property tends to make ensemble probabilities less extreme and, since the constituent models are overconfident, this results in better calibration (see also the histogram in \cref{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model}).

\begin{figure}
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_curves_ensemble_with_cis.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/calibration_fits_ensemble.pdf}
    \end{subfigure}    
    \caption[Calibration fits and curves for the stroke recognition ensemble using Platt-scaling and isotonic regression for calibration.]{ Calibration curves using sigmoid and isotonic calibration fits for the stroke recognition ensemble model (left) and the calibration fits (right) for the test set. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-calibration-curve-sigmoid-isotonic}
\end{figure}

To calibrate the ensemble model, we can use methods such as Platt-scaling \parencite{platt_probabilistic_1999} or isotonic regression \parencite{zadrozny_transforming_2002}. 
In either case, we fit a simple regression model (logistic or isotonic) to the predicted probabilities and the target labels on the validation set and use it to adjust the probabilities predicted on the test set. We show the resulting calibration curves on the left in \cref{fig_discussion:retrospective-paper-calibration-curve-sigmoid-isotonic} and the logistic and isotonic fits on the right with 95\% bootstrap confidence intervals on the bin centers ($x$ error) and fraction of positives ($y$ error). 
% In \cref{fig_discussion:retrospective-paper-calibration-curve-sigmoid-isotonic} we have done so for the ensemble model and visualize the results on the test set. 
% On the left, we plot the resulting calibration curves and on the right we show the logistic and isotonic fits. 
We see that both methods result in quite good calibrations\footnote{Brier scores on test set: Uncalibrated = $0.003500$, logistic = $0.001807$, isotonic = $0.001774$. Relative improvement in Brier score compared to uncalibrated (Brier skill score): Logistic = $0.4830$, isotonic = $0.4924$.} and that the predicted probabilities are shifted towards smaller values. 
Since stroke cases have low prevalence, high probability is predicted only for a few examples (see histogram in \cref{fig_discussion:retrospective-paper-calibration-curve-of-uncalibrated-model}). This leads to a lack of data for the calibration fits at high predicted probabilities which can be seen to result in poor generalization to the test set, especially for the nonparametric isotonic regression.

% Brier scores:  {'uncalibrated': 0.0034955788687128925, 'logistic': 0.0018072483306197486, 'isotonic': 0.0017744177396100578, 'average': 0.0021955477869598783}
% BSS uncalibrated reference:  {'logistic': 0.48299025755205005, 'isotonic': 0.4923822902432705}
% BSS mean reference:  {'logistic': 0.17685766561145932, 'isotonic': 0.1918109229282361}

\begin{figure}
    \centering
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/f1_score_vs_predicted_probability_ensemble_calltaker_with_cis.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[c]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/predicted_probability_histogram.pdf}
    \end{subfigure}    
    \caption[Comparison of F1-score of stroke recognition ensemble and call-takers as function of predicted probability.]{ Comparison of the F1-score of the stroke recognition ensemble and call-takers. The F1-score is computed on subsets of the test dataset made by binning on the predicted probabilities of the calibrated ensemble. We see that the relative performance improvement of the ensemble over call-takers is higher towards more certain predictions. We use the ensemble that achieved the median F1-score reported in \cref{fig_retrospective:figure1-roc-curve} and \cref{tab_retrospective:table3-occlusion-analysis}.}
    \label{fig_discussion:retrospective-paper-f1-performance-vs-predicted-probability}
    % \begin{subfigure}[c]{0.49\columnwidth}
    %     \centering
    %     \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/precision_vs_predicted_probability_ensemble_calltaker_with_cis.pdf}
    %     % \caption{}
    %     % \label{fig_discussion:calibration_curve_ensemble_and_all_models_uncalibrated}
    % \end{subfigure}    
    % % \hfill
    % \begin{subfigure}[c]{0.49\columnwidth}
    %     \centering
    %     \includegraphics[width=1\columnwidth]{paper_retrospective_calibration_plots/recall_vs_predicted_probability_ensemble_calltaker_with_cis.pdf}
    %     % \caption{}
    %     % \label{fig_discussion:histogram_ensemble_and_single_model}
    % \end{subfigure}
\end{figure}

Clinicians in intensive care units and emergency departments have been found to strongly agree that a singular focus on overall accuracy cannot alone ensure sustained trust in a model \cite{tonekaboni_what_2019}. Clinicians expect an alert to present a prediction that aligns with patient status. Despite expert-agreed thresholds for when alerts should be triggered, however, many alerts may not be aligned since class imbalance and ambiguous information in many predictive problems in healthcare can lead to models with relatively low predictive precision \parencite{umscheid_development_2015, cite14, cite15, wenstrup_retrospective_2023}. In turn, this is likely to lead to alarm fatigue \parencite{embi_evaluating_2012} and can undermine the sustained use and endorsement by clinicians of such systems \parencite{guidi_clinician_2015}. 
The stroke recognition ensemble presented in \cref{chp:paper-retrospective} is not exempt from this risk. With a precision of $24.9\%$ (95\% confidence interval $24.3-25.5\%$) it will on average be wrong three out of four times it predicts a stroke. This unfortunately risks alarm fatigue among its potential users diminishing its effect in practice.

% If the stroke recognition ensemble were to be deployed in a randomized controlled trial, s
Calibrated probabilities in the alerts presented to users might be a way to alleviate alarm fatigue. Calibrated probabilities would allow users to discern between certain and uncertain predictions and also enable the system to present users only with predictions that have at least a certain probability of being correct. Similar approaches have been suggested by clinicians and interviews indicate that predictive uncertainty is perceived by experts as a sort of explanation that complements the prediction \cite{tonekaboni_what_2019}. In \cref{fig_discussion:retrospective-paper-f1-performance-vs-predicted-probability} we show the F1-score of the ensemble model and the call-takers computed on subsets of the data created by binning the calibrated probabilities. We note that, as might be expected, both call-taker and ensemble model performance increase with increased model certainty. This indicates that selecting, based on certainty, which predictions to present to users might indeed help build trust in the system and ensure a practical impact.

% \lesstodo[inline]{Maybe mention MultiQT paper \parencite{havtorn_multiqt_2020}.}