### 2023-11-27 frontmatter/acknowledgements.tex
Write acknowledgments

%!TEX root = ../thesis.tex

\chapter[acknowledgements]{Acknowledgements}
First, I want to thank my supervisors. Jes, you have been an amazing support through this project. 
Your positivity and insightfulness were essential to me daring to take on this project, and to its success. 
Lars, you have been an indispensable source of purpose, energy, and direction. I look forward to many more fruitful discussions in the future.

I also want to thank the brilliant researchers at the section for Cognitive Systems that I have had the privilege of working with. 
Søren, as my co-supervisor, you have been a fantastic discussion partner. Your wit and your ability to pose genuinely significant questions have been invaluable. 
I want to also convey appreciation for the researchers in my group, for the open, inclusive, and relaxed culture, and for many good discussions, often over coffee and pastries.

I want to thank Lars, Andreas, and Michael for letting me become part of the Corti project after graduating back in 2018. 
The foundational will to invest and believe in people, and to dare to have bold visions, has made Corti a fantastic place for me to develop, professionally and personally. 
I want to also thank the members of the machine learning team during the first years of my time at Corti. Lasse, Tycho, Marco, Jan, Janek, and Joakim, you all played a part in inspiring and enabling me to dive into this project. 
A special thanks should go to Joakim. Your honest curiosity and dedication has been a wonderful addition to the research team. 
Finally, I want to thank Lasse. Our numerous inspirational discussions and combined efforts through the years have been instrumental in shaping the trajectory of this thesis. It has been a privilege to carry out this PhD project in the company of you both.

I want to also thank my friends and my family who have provided much needed diversion throughout the years. 
Last, but not least, I want to thank my wife, Rikke, who has been an unbelievably patient and absolutely indispensable supporter at every step of this undertaking.

\vspace*{\fill}
\noindent \textit{The project was funded by Corti and Innovation Fund Denmark through industrial PhD grant no.\@ 0153-00167B.}