### 2024-02-21 presentation/parts/overview.tex
WIP

% !TEX root = ../presentation.tex
% !BIB program = biber
% !TEX program = xelatex

\begin{frame}
    \frametitle{Overview: Thesis}
    \begin{columns}
        \begin{column}{0.6\textwidth}
            \begin{table}
                \resizebox*{\textwidth}{!}{%
                \begin{tabular}{l l}
                    \textsc{chapter 1-3}                   & \makecell[l]{Introduction, Research Questions, and Background} \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]
                    \textsc{chapter 4}                     & \makecell[l]{Hierarchical VAEs Know What They Don't Know} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    \makecell[l]{\textsc{chapter 5}\\}     & \makecell[l]{Model-Agnostic Out-of-Distribution Detection\\ Using Combined Statistical Tests} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    \makecell[l]{\textsc{chapter 6}\\}     & \makecell[l]{A Brief Overview of Unsupervised Speech \\Representation Learning} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    \textsc{chapter 7}     & \makecell[l]{Benchmarking Latent Variable Models for Speech} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    \makecell[l]{\textsc{chapter 8}\\}     & \makecell[l]{Automated Medical Coding on MIMIC-III and \\MIMIC-IV: A Critical Review and Replicability Study} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    \makecell[l]{\textsc{chapter 9}\\}     & \makecell[l]{A Retrospective Study on Machine Learning-\\Assisted Stroke Recognition for Medical Helpline Calls} \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]
                    \textsc{chapter 10}                    & \makecell[l]{Discussion and Conclusion} \\
                \end{tabular}
                }
            \end{table}
        \end{column}
        \begin{column}{0.4\textwidth}
        \end{column}
    \end{columns}
\end{frame}