### 2024-02-22 presentation/parts/overview.tex
WIP

% !TEX root = ../presentation.tex
% !BIB program = biber
% !TEX program = xelatex

\begin{frame}
    \frametitle{Overview: Thesis}
    \begin{columns}
        \begin{column}{0.6\textwidth}
            \begin{table}
                \resizebox*{\textwidth}{!}{%
                \begin{tabular}{l l l l}

                    & {\color<2->{black}   \makecell[l]{\textsc{\itshape chapter 1-3}}}       & {\color<2->{black}    \bfseries \scshape \Large \makecell[l]{introduction, research questions, and background}} & \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 4}}}         & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{hierarchical vaes know what they don't know}} & \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 5}\\\\}}     & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{model-agnostic out-of-distribution detection\\ using combined statistical tests}} & \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{ \textsc{\itshape chapter 6}\\\\}}    & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{a brief overview of unsupervised speech \\representation learning}} & \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 7}}}         & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{benchmarking latent variable models for speech}} & \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 8}\\\\}}     & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{automated medical coding on mimic-iii and \\mimic-iv: a critical review and replicability study}} & \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 9}\\\\}}     & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{a retrospective study on machine learning-\\assisted stroke recognition for medical helpline calls}} & \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 10}}}        & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{discussion and conclusion}} & \\
                \end{tabular}
                }
            \end{table}
        \end{column}
        \begin{column}{0.4\textwidth}
        \end{column}
    \end{columns}
\end{frame}