### 2024-02-22 presentation/parts/selection.tex

% !TEX root = ../presentation.tex
% !BIB program = biber
% !TEX program = xelatex

\begin{frame}
    \frametitle{Overview: Thesis}
    \begin{columns}
        \begin{column}{0.6\textwidth}

            \hfsetfillcolor{presentation_green}
            \hfsetbordercolor{presentation_green}

            \begin{table}
                \resizebox*{\textwidth}{!}{%
                \begin{tabular}{l l l l}

                    & {\color<2->{black!20} \textsc{chapter 1-3}}   & {\color<2->{black!20}\makecell[l]{Introduction, Research Questions, and Background}} & \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    \tikzmarkin<2>{hierarchical-1}(0.1,-0.3)(-0.1,0.4)
                    & \textsc{chapter 4}     & Hierarchical VAEs Know What They Don't Know & \tikzmarkend{hierarchical-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<3>{modelagnostic-1}(0.1,-0.4)(-0.1,0.6)
                    & \makecell[l]{\textsc{chapter 5}\\\\}     & \makecell[l]   {Model-Agnostic Out-of-Distribution Detection\\ Using Combined Statistical Tests} & \tikzmarkend{modelagnostic-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<4>{brief-1}(0.1,-0.4)(-0.1,0.6)
                    & \makecell[l]{\textsc{chapter 6}\\\\}     & \makecell[l]{A Brief Overview of Unsupervised Speech \\Representation Learning} & \tikzmarkend{brief-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<5>{benchmarking-1}(0.1,-0.3)(-0.1,0.4)
                    & \textsc{chapter 7}     & \makecell[l]{Benchmarking Latent Variable Models for Speech} & \tikzmarkend{benchmarking-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<6>{automated-1}(0.1,-0.4)(-0.1,0.6)
                    & \makecell[l]{\textsc{chapter 8}\\\\}     & \makecell[l]{Automated Medical Coding on MIMIC-III and \\MIMIC-IV: A Critical Review and Replicability Study} & \tikzmarkend{automated-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<7>{retrospective-1}(0.1,-0.4)(-0.1,0.6)
                    & \makecell[l]{\textsc{chapter 9}\\\\}     & \makecell[l]{A Retrospective Study on Machine Learning-\\Assisted Stroke Recognition for Medical Helpline Calls} & \tikzmarkend{retrospective-1} \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    & {\color<2->{black!20} \textsc{chapter 10}}    & {\color<2->{black!20} \makecell[l]{Discussion and Conclusion}} & \\
                \end{tabular}
                }
            \end{table}
        \end{column}
        \begin{column}{0.4\textwidth}

            \tikzstyle{mybox} = [fill, rounded corners, minimum width=140, minimum height=20]
            \tikzstyle{myrect} = [fill]

            \begin{figure}
                \centering
                \begin{overprint}
                    \onslide<1>  % Overview
                    \begin{tikzpicture}

                        \node [mybox, fill=presentation_orange] at (0,5) (predictions) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,5) {\small\textsc{predictions}};

                        \node [mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

                        \node [mybox, fill=presentation_blue] at (0,3) (text) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,3) {\small\textsc{text}};

                        \node [mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

                        \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

                        \node [mybox, fill=presentation_green] at (0,0) (speech) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,0) {\small\textsc{speech}};

                        \node [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
                        \node [above] at (uncertainty.north) {\small\textsc{uncertainty}};

                    \end{tikzpicture}

                    \onslide<2>  % Hierarchical
                    \begin{tikzpicture}[x=1cm,y=1cm]
                        \node [nearly transparent, mybox, fill=presentation_orange] at (0,5) (predictions) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,5) {\small\textsc{predictions}};

                        \node [nearly transparent, mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

                        \node [nearly transparent, mybox, fill=presentation_blue] at (0,3) (text) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,3) {\small\textsc{text}};

                        \node [nearly transparent, mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

                        \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

                        \node [nearly transparent, mybox, fill=presentation_green] at (0,0) (speech) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,0) {\small\textsc{speech}};

                        \node [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
                        \node [above] at (uncertainty.north) {\small\textsc{uncertainty}};
                    \end{tikzpicture}

                    \onslide<3>  % Model-Agnostic
                    \begin{tikzpicture}[x=1cm,y=1cm]
                        \node [nearly transparent, mybox, fill=presentation_orange] at (0,5) (predictions) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,5) {\small\textsc{predictions}};

                        \node [nearly transparent, mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

                        \node [nearly transparent, mybox, fill=presentation_blue] at (0,3) (text) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,3) {\small\textsc{text}};

                        \node [nearly transparent, mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

                        \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

                        \node [nearly transparent, mybox, fill=presentation_green] at (0,0) (speech) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,0) {\small\textsc{speech}};

                        \node [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
                        \node [above] at (uncertainty.north) {\small\textsc{uncertainty}};
                    \end{tikzpicture}

                    \onslide<4>  % Brief
                    \begin{tikzpicture}[x=1cm,y=1cm]
                        \node [nearly transparent, mybox, fill=presentation_orange] at (0,5) (predictions) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,5) {\small\textsc{predictions}};

                        \node [nearly transparent, mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

                        \node [nearly transparent, mybox, fill=presentation_blue] at (0,3) (text) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,3) {\small\textsc{text}};

                        \node [nearly transparent, mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

                        \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

                        \node [mybox, fill=presentation_green] at (0,0) (speech) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,0) {\small\textsc{speech}};

                        \node [nearly transparent, draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
                        \node [nearly transparent, above] at (uncertainty.north) {\small\textsc{uncertainty}};
                    \end{tikzpicture}

                    \onslide<5>  % Benchmarking
                    \begin{tikzpicture}[x=1cm,y=1cm]
                        \node [nearly transparent, mybox, fill=presentation_orange] at (0,5) (predictions) {};
                        \foreach \i in {-11,...,11} {
                            \fill[nearly transparent, myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
                        }
                        \node [nearly transparent] at (0,5) {\small\textsc{predictions}};

                        \node [nearly transparent, mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

                        \node [mybox, fill=presentation_blue] at (0,3) (text) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,3) {\small\textsc{text}};

                        \node [mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

                        \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

                        \node [mybox, fill=presentation_green] at (0,0) (speech) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,0) {\small\textsc{speech}};

                        \node [nearly transparent, draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
                        \node [nearly transparent, above] at (uncertainty.north) {\small\textsc{uncertainty}};
                    \end{tikzpicture}

                    \onslide<6>  % Automated
                    \begin{tikzpicture}[x=1cm,y=1cm]
                        \node [mybox, fill=presentation_orange] at (0,5) (predictions) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,5) {\small\textsc{predictions}};

                        \node [mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

                        \node [mybox, fill=presentation_blue] at (0,3) (text) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,3) {\small\textsc{text}};

                        \node [mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

                        \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

                        \node [mybox, fill=presentation_green] at (0,0) (speech) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,0) {\small\textsc{speech}};

                        \node [nearly transparent, draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
                        \node [nearly transparent, above] at (uncertainty.north) {\small\textsc{uncertainty}};
                    \end{tikzpicture}

                    \onslide<7>  % Retrospective
                    \begin{tikzpicture}[x=1cm,y=1cm]
                        \node [mybox, fill=presentation_orange] at (0,5) (predictions) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,5) {\small\textsc{predictions}};

                        \node [mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

                        \node [mybox, fill=presentation_blue] at (0,3) (text) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,3) {\small\textsc{text}};

                        \node [mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

                        \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

                        \node [mybox, fill=presentation_green] at (0,0) (speech) {};
                        \foreach \i in {-11,...,11} {
                            \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
                        }
                        \node [] at (0,0) {\small\textsc{speech}};

                        \node [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
                        \node [above] at (uncertainty.north) {\small\textsc{uncertainty}};
                    \end{tikzpicture}

                \end{overprint}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Overview: Presentation}
    \begin{columns}
        \begin{column}{0.6\textwidth}

            \hfsetfillcolor{presentation_green}
            \hfsetbordercolor{presentation_green}

            \begin{table}
                \resizebox*{\textwidth}{!}{%
                \begin{tabular}{l l l l}
                    & \textsc{chapter 1-3}   & {\color<2->{black!50}}\makecell[l]{Introduction, Research Questions, and Background} & \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]
                    \tikzmarkin<2>[below right offset={0.1,-0.4},above left offset={-0.1,0.4}]{a}
                    & \textsc{chapter 4}     & \makecell[l]{Hierarchical VAEs Know What They Don't Know} & \tikzmarkend{a}\\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    % \tikzmarkin<3>[below right offset={0.1,-0.6},above left offset={-0.1,0.6}]{b}
                    % & \makecell[l]{\textsc{chapter 5}}     & \makecell[l]{Model-Agnostic Out-of-Distribution Detection\\ Using Combined Statistical Tests} & \tikzmarkend{b} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    \tikzmarkin<4>[below right offset={0.1,-0.6},above left offset={-0.1,0.6}]{c}
                    & \makecell[l]{\textsc{chapter 6}}     & \makecell[l]{A Brief Overview of Unsupervised Speech \\Representation Learning} & \tikzmarkend{c} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    % \tikzmarkin<5>[below right offset={0.1,-0.4},above left offset={-0.1,0.4}]{d}
                    % & \textsc{chapter 7}     & \makecell[l]{Benchmarking Latent Variable Models for Speech} & \tikzmarkend{d} \\
                    % \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    % \tikzmarkin<6>[below right offset={0.1,-0.6},above left offset={-0.1,0.6}]{e}
                    % & \makecell[l]{\textsc{chapter 8}}     & \makecell[l]{Automated Medical Coding on MIMIC-III and \\MIMIC-IV: A Critical Review and Replicability Study} & \tikzmarkend{e} \\
                    % \addlinespace[0.5em]
                    \addlinespace[0.5em]
                    \tikzmarkin<7>[below right offset={0.1,-0.6},above left offset={-0.1,0.6}]{f}
                    & \makecell[l]{\textsc{chapter 9}}     & \makecell[l]{A Retrospective Study on Machine Learning-\\Assisted Stroke Recognition for Medical Helpline Calls} & \tikzmarkend{f} \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]
                    & \textsc{chapter 10}    & \makecell[l]{Discussion and Conclusion} & \\
                \end{tabular}
                }
            \end{table}
        \end{column}
    \end{columns}
\end{frame}