### 2024-02-25 presentation/parts/overview_brief.tex

% !TEX root = ../presentation.tex
% !BIB program = biber
% !TEX program = xelatex

\section{Overview}

\begin{frame}
    \frametitle{Presentation}
    \begin{columns}
        \begin{column}{0.6\textwidth}
            \begin{table}
                \resizebox{\textwidth}{!}{%
                \begin{tabular}{l l l l}
                    & {\color<1->{black!20}   \makecell[l]{\textsc{\itshape chapter 1-3}}} & {\color<1->{black!20}    \bfseries \scshape \Large \makecell[l]{introduction, research questions, and background}} & \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    \tikzmarkin<1->{hierarchical-7}(0.1,-0.3)(-0.1,0.4)
                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 4}}}         & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{hierarchical vaes know what they don't know}} & \tikzmarkend{hierarchical-7} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    % \tikzmarkin<1->{modelagnostic-7}(0.1,-0.4)(-0.1,0.6)
                    % & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 5}\\\\}}     & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{model-agnostic out-of-distribution detection\\ using combined statistical tests}} & \tikzmarkend{modelagnostic-7} \\
                    % \addlinespace[0.5em]
                    % \addlinespace[0.5em]

                    \tikzmarkin<1->{brief-7}(0.1,-0.4)(-0.1,0.6)
                    & {\color<-1>{black!20}\makecell[l]{ \textsc{\itshape chapter 6}\\\\}}    & {\color<-1>{black!20} \bfseries \scshape \Large \makecell[l]{a brief overview of unsupervised speech \\representation learning}} & \tikzmarkend{brief-7} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    % \tikzmarkin<1->{benchmarking-7}(0.1,-0.3)(-0.1,0.4)
                    % & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 7}}}         & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{benchmarking latent variable models for speech}} & \tikzmarkend{benchmarking-7} \\
                    % \addlinespace[0.5em]
                    % \addlinespace[0.5em]

                    % \tikzmarkin<1->{automated-7}(0.1,-0.4)(-0.1,0.6)
                    % & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 8}\\\\}}     & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{automated medical coding on mimic-iii and \\mimic-iv: a critical review and replicability study}} & \tikzmarkend{automated-7} \\
                    % \addlinespace[0.5em]
                    % \addlinespace[0.5em]

                    \tikzmarkin<1->{retrospective-7}(0.1,-0.4)(-0.1,0.6)
                    & {\color<1->{black!20}\makecell[l]{\textsc{\itshape chapter 9}\\\\}}     & {\color<1->{black!20} \bfseries \scshape \Large \makecell[l]{a retrospective study on machine learning-\\assisted stroke recognition for medical helpline calls}} & \tikzmarkend{retrospective-7} \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    & {\color<1->{black!20}\makecell[l]{\textsc{\itshape chapter 10}}}        & {\color<1->{black!20} \bfseries \scshape \Large \makecell[l]{discussion and conclusion}} & \\
                \end{tabular}
                }
            \end{table}
        \end{column}
        \begin{column}{0.4\textwidth}
            \tikzstyle{mybox} = [fill, rounded corners, minimum width=140, minimum height=20]
            \tikzstyle{myrect} = [fill]
            \begin{figure}
                \centering
                \begin{overprint}
                    \onslide<1>\tikzhierarchical
                    \onslide<2>\tikzbrief
                \end{overprint}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}