### 2024-02-27 presentation/parts/introduction.tex
WIP

% !TEX root = ../presentation.tex
% !BIB program = biber
% !TEX program = xelatex

\section{Introduction}

\begin{frame}
    % Background on Corti and the problem of medical communication, diagnosis, and documentation.
    \frametitle{Background}
    \begin{columns}
        \begin{column}{0.6\textwidth}
            FILL ME OUT. \\
            Background on Corti, motivation for project.
        \end{column}
        \begin{column}{0.4\textwidth}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Healthcare}
    \begin{quotation}
        \noindent\centering
        Healthcare is the improvement of health via the \highlight{prevention}, \highlight{diagnosis}, \highlight{treatment}, \highlight{amelioration} or \highlight{cure} of {\color{theme-green}disease}, {\color{theme-green}illness}, {\color{theme-green}injury}, and {\color{theme-green}other physical and mental impairments} in people.
    \end{quotation}
\end{frame}

% \begin{frame}
%     \frametitle{Medical dialogue}
%     \begin{figure}
%         \centering
%         \includegraphics[width=0.7\paperwidth]{figures/speech_bubble.pdf}
%     \end{figure}

%     \note[item]{Healthcare sector is a complex system}
% \end{frame}
% \begin{frame}
%     \frametitle{Medical dialogue}
%     \begin{figure}
%         \centering
%         \includegraphics[width=0.95\textheight]{figures/corti_sketch_conversations.png}
%     \end{figure}
% \end{frame}

\begin{frame}[t]
    \frametitle{Medical dialogue}
    \begin{columns}[t]
        \begin{column}{0.4\textwidth}
            \begin{itemize}
                \item <2> General practitioner
                \item <2> Nurse
                \item <2> Midwife
                \item <2> Emergency medical dispatcher
                \item <2> Paramedic
                \item <2> Emergency room
                \item <2> Health insurance
            \end{itemize}
        \end{column}
        \begin{column}{0.6\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=0.9\textwidth]{figures/corti_sketch_conversations.png}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Errors in medical dialogue}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item <1-> Communication is everywhere in healthcare. 
                \item <1-> It is complex, involving multiple participants, different contexts, and different purposes.
                \vspace{1em}
                \item <2-> \highlight{Adverse events}: Failure of communication is a leading cause of medical error contributing to two out of three adverse events \cite{starmer_changes_2014}.
                \item <2-> \highlight{Preventability}: A considerable fraction of all hospital admissions have preventable adverse outcomes\footnote<2>{9\% to 16.6\% in AU, NZ, UK, DK.} \cite{carver_medical_2024}.
            \end{itemize}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=0.95\textwidth]{figures/corti_sketch_emergency.png}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Documenting medical encounters}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item <1,2> Documentation is a central part of healthcare.
                \item <1,2> E.g. patient records, insurance claims, billing, research, training, legal purposes.
                \vspace{1em}
                \item <2> \highlight{Time-consuming}: Physicians spend 34-37\% of their time on documentation \cite{joukes_time_2018, tipping_where_2010, sinsky_allocation_2016}\footnote<2>{Ambulatory care across four specialties in four states and tertiary care at an academic medical center.}.
                \item <2> \highlight{Varying quality}: Discharge summaries almost never meet \emph{all} timeline, transmission, and content criteria. \cite{horwitz_comprehensive_2013}\footnote<2>{Outpatient visits, Yale-New Haven Hospital.}
            \end{itemize}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=0.95\textwidth]{figures/corti_sketch_responders.png}
            \end{figure}
        \end{column}
    \end{columns}
    \note[item]{Documentation is essential for a number of purposes but is very time-consuming and imperfect}
\end{frame}

\begin{frame}
    \frametitle{How might machine learning help?}
    \begin{columns}
        \begin{column}{0.35\textwidth}
            \begin{itemize}
                \item <1-> \highlight{Assist} with documentation.
                \item <1-> \highlight{Augment} communication.
                \item <1-> \highlight{Improve} decision-making.
                \vspace{1em}
                \item <2-> \highlight{Reduce} number of impact of medical errors and adverse events.
                \item <2-> \highlight{Free up} time spent on documentation for patient care.
            \end{itemize}
        \end{column}
        \begin{column}{0.65\textwidth}
            \centering
            \includegraphics[width=0.95\textwidth]{figures/corti_sketch_office.png}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Reliability of machine learning systems}
    \begin{itemize}
        \item <1-> \highlight{Data}: Quality, quantity, diversity, bias, privacy, ethics.
        \item <2-> \highlight{Task}: Context, domain, language, culture, purpose.
        \vspace{1em}
        \item <3-> \highlight{Interpretability} of how a model works (transparency, accountability, regulation).
        \item <4-> \highlight{Explainability} of model predictions (trust, understanding, feedback).
        \item <5-> \highlight{Fairness} in treatment of different groups of people.
        \item <6-> \highlight{Robustness} to noise, outliers, adversarial attacks, and distribution shift.
    \end{itemize}
\end{frame}