### 2024-03-02 presentation/parts/titlepage.tex

% !TEX root = presentation.tex
% !BIB program = biber
% !TEX program = xelatex

\frame{
    \maketitle

    \note{
        \begin{itemize}
            \item Welcome to my PhD defense.
            \item Thank you to the moderator and the assessment committee for taking part today.
            \item I will present my work on uncertainty estimation in AI systems for medical domains.
            \item I will start with an overview of the thesis followed by a brief introduction.
            \item Then I will present a selection of the research chapters.
            \item Finally, I will discuss the broader implications of the work.
        \end{itemize}
    }
}