### 2024-03-03 presentation/parts/overview_presentation.tex
WIP

% !TEX root = ../presentation.tex
% !BIB program = biber
% !TEX program = xelatex

\hfsetfillcolor{presentation_green}
\hfsetbordercolor{presentation_green}

\newcommand{\tikzoverview}{%
    \resizebox{0.8\textwidth}{!}{%
        \begin{tikzpicture}[x=1cm,y=1cm]
            \node [mybox, fill=presentation_orange] at (0,5) (predictions) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
            }
            \node [] at (0,5) {\small\textsc{predictions}};

            \node [mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

            \node [mybox, fill=presentation_blue] at (0,3) (text) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
            }
            \node [] at (0,3) {\small\textsc{text}};

            \node [mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

            \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

            \node [mybox, fill=presentation_green] at (0,0) (speech) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
            }
            \node [] at (0,0) {\small\textsc{speech}};

            \node [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
            \node [above] at (uncertainty.north) {\small\textsc{uncertainty}};
        \end{tikzpicture}
    }
}

\newcommand{\tikzhierarchical}{%
    \resizebox{0.8\textwidth}{!}{%
        \begin{tikzpicture}[x=1cm,y=1cm]
            \node [nearly transparent, mybox, fill=presentation_orange] at (0,5) (predictions) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,5) {\small\textsc{predictions}};

            \node [nearly transparent, mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

            \node [nearly transparent, mybox, fill=presentation_blue] at (0,3) (text) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,3) {\small\textsc{text}};

            \node [nearly transparent, mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

            \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

            \node [nearly transparent, mybox, fill=presentation_green] at (0,0) (speech) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,0) {\small\textsc{speech}};

            \node [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
            \node [above] at (uncertainty.north) {\small\textsc{uncertainty}};
        \end{tikzpicture}
    }
}

\newcommand{\tikzmodelagnostic}{%
    \resizebox{0.8\textwidth}{!}{%
        \begin{tikzpicture}[x=1cm,y=1cm]
            \node [nearly transparent, mybox, fill=presentation_orange] at (0,5) (predictions) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,5) {\small\textsc{predictions}};

            \node [nearly transparent, mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

            \node [nearly transparent, mybox, fill=presentation_blue] at (0,3) (text) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,3) {\small\textsc{text}};

            \node [nearly transparent, mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

            \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

            \node [nearly transparent, mybox, fill=presentation_green] at (0,0) (speech) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,0) {\small\textsc{speech}};

            \node [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
            \node [above] at (uncertainty.north) {\small\textsc{uncertainty}};
        \end{tikzpicture}
    }
}

\newcommand{\tikzbrief}{%
    \resizebox{0.8\textwidth}{!}{%
        \begin{tikzpicture}[x=1cm,y=1cm]
            \node [nearly transparent, mybox, fill=presentation_orange] at (0,5) (predictions) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,5) {\small\textsc{predictions}};

            \node [nearly transparent, mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

            \node [nearly transparent, mybox, fill=presentation_blue] at (0,3) (text) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,3) {\small\textsc{text}};

            \node [nearly transparent, mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

            \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

            \node [mybox, fill=presentation_green] at (0,0) (speech) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
            }
            \node [] at (0,0) {\small\textsc{speech}};

            \node [nearly transparent, draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
            \node [nearly transparent, above] at (uncertainty.north) {\small\textsc{uncertainty}};
        \end{tikzpicture}
    }
}

\newcommand{\tikzbenchmarking}{%
    \resizebox{0.8\textwidth}{!}{%
        \begin{tikzpicture}[x=1cm,y=1cm]
            \node [nearly transparent, mybox, fill=presentation_orange] at (0,5) (predictions) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,5) {\small\textsc{predictions}};

            \node [nearly transparent, mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

            \node [mybox, fill=presentation_blue] at (0,3) (text) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
            }
            \node [] at (0,3) {\small\textsc{text}};

            \node [mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

            \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

            \node [mybox, fill=presentation_green] at (0,0) (speech) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
            }
            \node [] at (0,0) {\small\textsc{speech}};

            \node [nearly transparent, draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
            \node [nearly transparent, above] at (uncertainty.north) {\small\textsc{uncertainty}};
        \end{tikzpicture}
    }
}

\newcommand{\tikzautomated}{%
    \resizebox{0.8\textwidth}{!}{%
        \begin{tikzpicture}[x=1cm,y=1cm]
            \node [mybox, fill=presentation_orange] at (0,5) (predictions) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
            }
            \node [] at (0,5) {\small\textsc{predictions}};

            \node [mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

            \node [mybox, fill=presentation_blue] at (0,3) (text) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
            }
            \node [] at (0,3) {\small\textsc{text}};

            \node [nearly transparent, mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

            \node [nearly transparent, mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

            \node [nearly transparent, mybox, fill=presentation_green] at (0,0) (speech) {};
            \foreach \i in {-11,...,11} {
                \fill[nearly transparent, myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
            }
            \node [nearly transparent] at (0,0) {\small\textsc{speech}};

            \node [nearly transparent, draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
            \node [nearly transparent, above] at (uncertainty.north) {\small\textsc{uncertainty}};
        \end{tikzpicture}
    }
}

\newcommand{\tikzretrospective}{%
    \resizebox{0.8\textwidth}{!}{%
        \begin{tikzpicture}[x=1cm,y=1cm]
            \node [mybox, fill=presentation_orange] at (0,5) (predictions) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
            }
            \node [] at (0,5) {\small\textsc{predictions}};

            \node [mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

            \node [mybox, fill=presentation_blue] at (0,3) (text) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
            }
            \node [] at (0,3) {\small\textsc{text}};

            \node [mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

            \node [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

            \node [mybox, fill=presentation_green] at (0,0) (speech) {};
            \foreach \i in {-11,...,11} {
                \fill[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
            }
            \node [] at (0,0) {\small\textsc{speech}};

            \node [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
            \node [above] at (uncertainty.north) {\small\textsc{uncertainty}};
        \end{tikzpicture}
    }
}

\section{Overview}

% Go through the papers of the thesis while showing the graphic of their themes.
\begin{frame}
    \frametitle{Thesis}
    \begin{columns}
        \begin{column}{0.6\textwidth}
            \begin{table}
                \resizebox{\textwidth}{!}{%
                \begin{tabular}{l l l l}
                    & {\color<2->{black!20}   \makecell[l]{\textsc{\itshape chapter 1-3}}} & {\color<2->{black!20}    \bfseries \scshape \Large \makecell[l]{introduction, research questions, and background}} & \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    \tikzmarkin<2>{hierarchical-1}(0.1,-0.3)(-0.1,0.4)
                    & {\color<2->{black}\makecell[l]{\textsc{\itshape chapter 4}}}         & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{hierarchical vaes know what they don't know}} & \tikzmarkend{hierarchical-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<2>{modelagnostic-1}(0.1,-0.4)(-0.1,0.6)
                    & {\color<2->{black}\makecell[l]{\textsc{\itshape chapter 5}\\\\}}     & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{model-agnostic out-of-distribution detection\\ using combined statistical tests}} & \tikzmarkend{modelagnostic-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<3>{brief-1}(0.1,-0.4)(-0.1,0.6)
                    & {\color<2->{black}\makecell[l]{ \textsc{\itshape chapter 6}\\\\}}    & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{a brief overview of unsupervised speech \\representation learning}} & \tikzmarkend{brief-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<4>{benchmarking-1}(0.1,-0.3)(-0.1,0.4)
                    & {\color<2->{black}\makecell[l]{\textsc{\itshape chapter 7}}}         & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{benchmarking latent variable models for speech}} & \tikzmarkend{benchmarking-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<5>{automated-1}(0.1,-0.4)(-0.1,0.6)
                    & {\color<2->{black}\makecell[l]{\textsc{\itshape chapter 8}\\\\}}     & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{automated medical coding on mimic-iii and \\mimic-iv: a critical review and replicability study}} & \tikzmarkend{automated-1} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<6>{retrospective-1}(0.1,-0.4)(-0.1,0.6)
                    & {\color<2->{black}\makecell[l]{\textsc{\itshape chapter 9}\\\\}}     & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{a retrospective study on machine learning-\\assisted stroke recognition for medical helpline calls}} & \tikzmarkend{retrospective-1} \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 10}}}        & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{discussion and conclusion}} & \\
                \end{tabular}
                }
            \end{table}
        \end{column}
        \begin{column}{0.4\textwidth}
            \begin{figure}
                \centering
                \begin{overprint}
                    \onslide<1>\tikzoverview
                    \onslide<2>\tikzhierarchical
                    % \onslide<2>\tikzmodelagnostic
                    \onslide<3>\tikzbrief
                    \onslide<4>\tikzbenchmarking
                    \onslide<5>\tikzautomated
                    \onslide<6>\tikzretrospective
                \end{overprint}
            \end{figure}
        \end{column}
    \end{columns}

    \note[item]<1->{So how does the thesis tackle this problem of building such a system?}
    \note[item]<2->{First we examine how a certain class of generative models can be used to detect data that is different from training data.}
    \note[item]<2->{In our second paper we generalize this to be model-agnostic and phrased as a statistical test.}
    \note[item]<3->{We then take a step back and provide an overview of modern methods for unsupervised speech representation learning.}
    \note[item]<4->{We then benchmark some of these methods, generative latent variable models, on likelihood and automatic speech recognition.}
    \note[item]<5->{We then move to a different domain, medical coding, and provide a critical review of recent works.}
    \note[item]<6->{Finally, we provide a retrospective study on a machine learning-assisted stroke recognition system for medical helpline calls.}
    \end{frame}

\hfsetfillcolor{white}
\hfsetbordercolor{white}

% Grey out scoped-out papers in the overview
\begin{frame}
    \frametitle{Presentation}
    \begin{columns}
        \begin{column}{0.6\textwidth}
            \begin{table}
                \resizebox{\textwidth}{!}{%
                \begin{tabular}{l l l l}
                    & {\color<2->{black}   \makecell[l]{\textsc{\itshape chapter 1-3}}} & {\color<2->{black}    \bfseries \scshape \Large \makecell[l]{introduction, research questions, and background}} & \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    \tikzmarkin<1->{hierarchical-2}(0.1,-0.3)(-0.1,0.4)
                    & {\color<2->{black}\makecell[l]{\textsc{\itshape chapter 4}}}         & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{hierarchical vaes know what they don't know}} & \tikzmarkend{hierarchical-2} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<1->{modelagnostic-2}(0.1,-0.4)(-0.1,0.6)
                    & {\color<2->{dtured!40}\makecell[l]{\textsc{\itshape chapter 5}\\\\}}     & {\color<2->{dtured!40} \bfseries \scshape \Large \makecell[l]{model-agnostic out-of-distribution detection\\ using combined statistical tests}} & \tikzmarkend{modelagnostic-2} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<1->{brief-2}(0.1,-0.4)(-0.1,0.6)
                    & {\color<2->{black}\makecell[l]{ \textsc{\itshape chapter 6}\\\\}}    & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{a brief overview of unsupervised speech \\representation learning}} & \tikzmarkend{brief-2} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<1->{benchmarking-2}(0.1,-0.3)(-0.1,0.4)
                    & {\color<2->{dtured!40}\makecell[l]{\textsc{\itshape chapter 7}}}         & {\color<2->{dtured!40} \bfseries \scshape \Large \makecell[l]{benchmarking latent variable models for speech}} & \tikzmarkend{benchmarking-2} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<1->{automated-2}(0.1,-0.4)(-0.1,0.6)
                    & {\color<2->{dtured!40}\makecell[l]{\textsc{\itshape chapter 8}\\\\}}     & {\color<2->{dtured!40} \bfseries \scshape \Large \makecell[l]{automated medical coding on mimic-iii and \\mimic-iv: a critical review and replicability study}} & \tikzmarkend{automated-2} \\
                    \addlinespace[0.5em]
                    \addlinespace[0.5em]

                    \tikzmarkin<1->{retrospective-2}(0.1,-0.4)(-0.1,0.6)
                    & {\color<2->{black}\makecell[l]{\textsc{\itshape chapter 9}\\\\}}     & {\color<2->{black} \bfseries \scshape \Large \makecell[l]{a retrospective study on machine learning-\\assisted stroke recognition for medical helpline calls}} & \tikzmarkend{retrospective-2} \\
                    \addlinespace[0.5em]
                    \midrule
                    \addlinespace[0.5em]

                    & {\color<2->{black!20}\makecell[l]{\textsc{\itshape chapter 10}}}        & {\color<2->{black!20} \bfseries \scshape \Large \makecell[l]{discussion and conclusion}} & \\
                \end{tabular}
                }
            \end{table}
        \end{column}
        \begin{column}{0.4\textwidth}
            \tikzstyle{mybox} = [fill, rounded corners, minimum width=140, minimum height=20]
            \tikzstyle{myrect} = [fill]
            \begin{figure}
                \centering
                \begin{overprint}
                    \onslide<1->\tikzoverview
                \end{overprint}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}