### 2024-03-03 presentation/parts/presentation_introduction.tex

% !TEX root = ../presentation.tex
% !BIB program = biber
% !TEX program = xelatex

\section{introduction}

\begin{frame}
    \frametitle{What is healthcare?}
    \begin{quotation}
        \noindent\centering
        \textbf{Healthcare} is the improvement of health via the \highlight{prevention}, \highlight{diagnosis}, \highlight{treatment}, \highlight{amelioration} or \highlight{cure} of {\color{theme-green}disease}, {\color{theme-green}illness}, {\color{theme-green}injury}, and {\color{theme-green}other physical and mental impairments} in people.
    \end{quotation}
    \begin{flushright}
        \footnotesize
        --- Wikipedia
    \end{flushright}

    \note[item]{Machine learning for healthcare touches on most of these aspects. Fx:}
    \note[item]{
        \begin{itemize}
            \item \textbf{Prevention} by predicting risk of disease allowing for early intervention.
            \item \textbf{Diagnosis} by detecting stroke in emergency calls.
            \item \textbf{Treatment} by suggesting medication.
        \end{itemize}
    }
\end{frame}

\begin{frame}[t]
    \frametitle{Medical dialogue}
    \begin{columns}[t]
        \begin{column}{0.4\textwidth}
            Central to an \highlight{interaction} within a healthcare system is the \highlight{medical dialogue}:
            % Every interaction within a healthcare system is a form of \highlight{medical dialogue}:
            \begin{itemize}
                \item <1> General practitioner
                \item <1> Nurse
                \item <1> Midwife
                \item <1> Emergency medical dispatcher
                \item <1> Paramedic
                \item <1> Emergency room
                \item <1> Health insurance
            \end{itemize}
        \end{column}
        \begin{column}{0.6\textwidth}
            \vspace{-2em}
            \begin{figure}[t]
                \centering
                \includegraphics[width=0.9\textwidth]{figures/corti_sketch_conversations.png}
            \end{figure}
        \end{column}
    \end{columns}

    \note[item]{We focus on medical communication.}
    \note[item]{Involves many different parties.}
    \note[item]{Different contexts and purposes.}
    \note[item]{Busy emergency room, calls to emergency medical dispatchers, visits at general practitioners, etc.}
\end{frame}

\begin{frame}
    \frametitle{Errors in medical dialogue}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item <1-> Communication is everywhere in healthcare. 
                \item <1-> It is complex, involving multiple participants, different contexts, and different purposes.
                \vspace{1em}
                \item <2-> \highlight{Adverse events}: Failure of communication contributes to two out of three adverse events \cite{starmer_changes_2014}.
                \item <2-> \highlight{Preventability}: Many adverse outcomes are preventable \cite{carver_medical_2024}.
                % \footnote<2>{9\% to 16.6\% in AU, NZ, UK, DK.} \cite{carver_medical_2024}.
            \end{itemize}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=0.95\textwidth]{figures/corti_sketch_emergency.png}
            \end{figure}
        \end{column}
    \end{columns}

    \note[item]{So yes communication can be complex and noisy, but what are the consequences?}
    \note[item]{Failure of communication is a leading cause of adverse events.}
    \note[item]{Adverse events are episodes of medical error that result in harm to the patient.}
    \note[item]{Luckily, many of these are preventable, although exact numbers vary.}
    \note[item]{Better communication could help reduce these numbers.}
\end{frame}

\begin{frame}
    \frametitle{Documenting medical encounters}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item <1-> Documentation is a central part of healthcare.
                \item <1-> E.g. patient records, insurance claims, billing, research, training, legal purposes.
                \vspace{1em}
                \item <2-> \highlight{Time-consuming}: Physicians spend 34-37\% of their time on documentation \cite{joukes_time_2018, tipping_where_2010, sinsky_allocation_2016}\footnote<2->{Ambulatory care across four specialties in four states and tertiary care at an academic medical center.}.
                \item <2-> \highlight{Varying quality}: Discharge summaries almost never meet \emph{all} timeline, transmission, and content criteria. \cite{horwitz_comprehensive_2013}\footnote<2->{Outpatient visits, Yale-New Haven Hospital.}
            \end{itemize}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=0.95\textwidth]{figures/corti_sketch_responders.png}
            \end{figure}
        \end{column}
    \end{columns}

    \note[item]{Another central aspect of medical communication is documentation.}
    \note[item]{Essential for a number of purposes}
    \note[item]{But, it is time-consuming and of varying quality.}
    \note[item]{(Ambulatory $\equiv$ outpatient care, Tertiary $\equiv$ specialized care)}
\end{frame}

\begin{frame}
    \frametitle{How might machine learning help?}
    \begin{columns}
        \begin{column}{0.35\textwidth}
            \begin{itemize}
                \item <1-> \highlight{Assist} with documentation.
                \item <1-> \highlight{Augment} communication.
                \item <1-> \highlight{Improve} decision-making.
                \vspace{1em}
                \item <2-> \highlight{Reduce} the impact of medical errors and adverse events.
                \item <2-> \highlight{Free up} time spent on documentation for patient care.
            \end{itemize}
        \end{column}
        \begin{column}{0.65\textwidth}
            \centering
            \includegraphics[width=0.95\textwidth]{figures/corti_sketch_office.png}
        \end{column}
    \end{columns}

    \note[item]{Machine learning can help in many ways.}
    \note[item]{The main goal of augmenting communication is:
        \begin{itemize}
            \item Reduce the impact.
            \item Free up time.
        \end{itemize}
    }
\end{frame}

% \begin{frame}
%     \frametitle{Reliability of machine learning systems}
%     \begin{itemize}
%         \item <1-> \highlight{Data}: Quality, quantity, diversity, bias, privacy, ethics.
%         \item <1-> \highlight{Task}: Context, domain, language, culture, purpose.
%         \vspace{1em}
%         \item <2-> \highlight{Interpretability} of how a model works (transparency, accountability, regulation).
%         \item <2-> \highlight{Explainability} of model predictions (trust, understanding, feedback).
%         \item <2-> \highlight{Fairness} in treatment of different groups of people.
%         \item <2-> \highlight{Robustness} to noise, outliers, distribution shift, and adversarial attacks.
%     \end{itemize}

%     \note[item]{So what are the challenges holding us back in implementing such systems?}
%     \note[item]{
%         \begin{itemize}
%             \item Strong requirements of data.
%             \item Tasks that span different contexts, domains, languages, and cultures.
%             \item Regulatory requirements for transparency and accountability.
%             \item Trust and understanding of predictions.
%         \end{itemize}
%     }
% \end{frame}

\tikzstyle{mybox} = [fill, rounded corners, minimum width=140, minimum height=20]
\tikzstyle{myrect} = [fill]

\begin{frame}
    \frametitle{Building a decision-support system}
    \begin{columns}
        \begin{column}{0.6\textwidth}
            % Modular approach:
            \begin{itemize}
                \item <2-> \highlight{Source data}: Speech or text (potentially images, video, electronic health records, etc.).
                \item <3-> \highlight{Foundation modelling}: Converting the input into representations useful for downstream tasks.
                \item <4-> \highlight{Downstream tasks}: Suggesting, summarizing, classifying, transcribing, translating, etc.
                \item <5-> \highlight{Uncertainty}: Estimating the reliability of data, representations, predictions.
            \end{itemize}
        \end{column}
        \begin{column}{0.4\textwidth}
            \resizebox{0.8\textwidth}{!}{%
            \begin{tikzpicture}[x=1cm,y=1cm]
                \node [mybox, fill=white] at (0,5) (predictions_white) {};
                \node<4-> [mybox, fill=presentation_orange] at (0,5) (predictions) {};
                \foreach \i in {-11,...,11} {
                    \fill<4->[myrect, fill=presentation_orange_shaded] (0.2*\i,4.8) rectangle ++(0.1,0.4);
                }
                \node<4-> [] at (0,5) {\small\textsc{predictions}};

                \node<4-> [mybox, fill=black!10] at (0,4) (downstream) {\small\textsc{downstream model}};

                \node [mybox, fill=white] at (0,3) (text_white) {};
                \node<2-> [mybox, fill=presentation_blue] at (0,3) (text) {};
                \foreach \i in {-11,...,11} {
                    \fill<2->[myrect, fill=presentation_blue_shaded] (0.2*\i,2.8) rectangle ++(0.1,0.4);
                }
                \node<2-> [] at (0,3) {\small\textsc{text}};

                \node [mybox, fill=white] at (0,2) (recognition_white) {};
                \node<3-> [mybox, fill=black!10] at (0,2) (recognition) {\small\textsc{speech recognition}};

                \node [mybox, fill=white] at (0,1) (representation_white) {};
                \node<3-> [mybox, fill=black!10] at (0,1) (representation) {\small\textsc{representation learning}};

                \node [mybox, fill=white] at (0,0) (speech_white) {};
                \node<2-> [mybox, fill=presentation_green] at (0,0) (speech) {};
                \foreach \i in {-11,...,11} {
                    \fill<2->[myrect, fill=presentation_green_shaded] (0.2*\i,-0.2) rectangle ++(0.1,0.4);
                }
                \node<2-> [] at (0,0) {\small\textsc{speech}};

                \node [draw, rectangle, rounded corners, line width=0.15cm, white, fit=(speech_white)(representation_white)(recognition_white)(text_white)(predictions_white), inner sep=10pt] (uncertainty_white) {};
                \node [above, white] at (uncertainty_white.north) {\color{white}\small\textsc{uncertainty}};

                \node<5-> [draw, rectangle, rounded corners, line width=0.15cm, presentation_red, fit=(speech)(representation)(recognition)(text)(predictions), inner sep=10pt] (uncertainty) {};
                \node<5-> [above] at (uncertainty.north) {\small\textsc{uncertainty}};
            \end{tikzpicture}
        }
        \end{column}
    \end{columns}

    \note[item]{We will take a modular approach to building a decision-support system.}
    \note[item]{First we need source data}
    \note[item]{Then we need to convert it into representations useful for downstream tasks.}
    \note[item]{Then we can perform the downstream tasks.}
    \note[item]{Finally, we need to estimate the reliability of our data, representations, and predictions.}
\end{frame}